---
title: "Data 621 Homework 1: Code Appendix"
author: "Jeff Nieman, Scott Karr, James Topor, Armenoush Aslanian-Persico"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
    theme: spacelab
---


# Part 1. Data Exploration

```{r part1-load-libraries, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
library(alr3)
library(car)
library(corrplot)
library(fBasics)
library(knitr)
library(MASS)
```


```{r, message=F, warning=F}
library(psych)
library(car)
library(corrplot)
```

Load and explore original data

```{r}
mb_train <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/moneyball-training-data.csv")
describe(mb_train)

boxplot(mb_train, xlab="Boxplot Predictor Comparitive") 

```

Look for correlations among values

```{r, echo=FALSE}
m <- na.omit(mb_train)
cor(m)

```


# Part 2. Data Preparation

## New Variable Creation

#####Creating a new column for batting singles and eliminating hits for batting
```{r part2, eval=FALSE}
mb_e <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/moneyball-training-data.csv") 
#eliminate index column
mb_e1 <- mb_e[,-1]

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-2]
mb_e1 <- as.data.frame(mb_e1)
```

#####Building a regression model and filling in NA's for SB.  
Note:  This approach is suggested in LMAR p. 201.  "A more sophisticated alternative to mean imputation is to use regression methods to predict the missing values of the covariates."

```{r mb_e1, eval=FALSE}
SB <- lm(data=mb_e1, TEAM_BASERUN_SB~.)
summary(SB)

#eliminate CS as there are no blank SB's with a value for CS + eliminate pitching, wins and fielding variables
SB1 <- lm(data=mb_e1, TEAM_BASERUN_SB~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB)
summary(SB1)

#eliminate singles
SB2 <- lm(data=mb_e1, TEAM_BASERUN_SB~TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB)
summary(SB2)

#fill in NA for SB
mb_e2 <- mb_e1
mb_e2$TEAM_BASERUN_SB[is.na(mb_e2$TEAM_BASERUN_SB)]<-round(79.36805-0.19419*mb_e2$TEAM_BATTING_2B+1.41686*mb_e2$TEAM_BATTING_3B-0.24513*mb_e2$TEAM_BATTING_HR+0.08060*mb_e2$TEAM_BATTING_BB)
```

#####Building a linear regression model and filling in NA's for CS
```{r mb_e2, eval=FALSE}
CS <- lm(data=mb_e2, TEAM_BASERUN_CS~.)
summary(CS)

#eliminate wins, pitching and fielding 
CS1 <- lm(data=mb_e2, TEAM_BASERUN_CS~TEAM_BASERUN_SB +TEAM_BATTING_1B +TEAM_BATTING_2B + TEAM_BATTING_3B +TEAM_BATTING_BB + TEAM_BATTING_HR)
summary(CS1)

#eliminate walks
CS2 <- lm(data=mb_e2, TEAM_BASERUN_CS~TEAM_BASERUN_SB +TEAM_BATTING_2B + TEAM_BATTING_3B +TEAM_BATTING_1B + TEAM_BATTING_HR)
summary(CS2)

#fill in NA for CS
mb_e3 <- mb_e2
mb_e3$TEAM_BASERUN_CS[is.na(mb_e3$TEAM_BASERUN_CS)]<-round(49.356793+0.322543*mb_e3$TEAM_BASERUN_SB-0.044486*mb_e3$TEAM_BATTING_2B+0.281124*mb_e3$TEAM_BATTING_3B-0.10797*mb_e3$TEAM_BATTING_HR-0.014034*mb_e3$TEAM_BATTING_1B)

```

#####Building a regression model and filling in NA's for batting SO's
```{r mb_e3, eval=FALSE}
BSO <- lm(data=mb_e3, TEAM_BATTING_SO~.)
summary(BSO)

#eliminate fielding and wins and baserunning and HBP and pitching SO's as it contains similar blanks
BSO1 <- lm(data=mb_e3, TEAM_BATTING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(BSO1)

#eliminate pitching HR's
BSO2 <- lm(data=mb_e3, TEAM_BATTING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_BB + TEAM_PITCHING_H  + TEAM_PITCHING_BB + TEAM_BATTING_HR)
summary(BSO2)

#fill in NA for batting SO
mb_e4 <- mb_e3
mb_e4$TEAM_BATTING_SO[is.na(mb_e4$TEAM_BATTING_SO)]<-round(1605-0.8434*mb_e4$TEAM_BATTING_1B+0.2832*mb_e4$TEAM_BATTING_2B-1.348*mb_e4$TEAM_BATTING_3B-0.3493*mb_e4$TEAM_BATTING_BB-0.02903*mb_e4$TEAM_PITCHING_H+0.1657*mb_e4$TEAM_PITCHING_BB+1.703*mb_e$TEAM_BATTING_HR)

```

#####Building a regression model and filling in NA's for pitching SO's
```{r mb_e4, eval=FALSE}
PSO <- lm(data=mb_e4, TEAM_PITCHING_SO~.)
summary(PSO)

#eliminate wins, fielding, baserunning and HBP
PSO1 <- lm(data=mb_e4, TEAM_PITCHING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_BATTING_SO + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(PSO1)

#eliminate batting 3B's
PSO2 <- lm(data=mb_e4, TEAM_PITCHING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_BATTING_SO + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(PSO2)

#replace NA with values for pitching SO
mb_e5 <- mb_e4
mb_e5$TEAM_PITCHING_SO[is.na(mb_e5$TEAM_PITCHING_SO)]<-round(4422.87422-0.46455*mb_e5$TEAM_BATTING_1B + 0.72823*mb_e5$TEAM_BATTING_2B +8.06941*mb_e5$TEAM_BATTING_HR -3.45005*mb_e5$TEAM_BATTING_BB-0.01815*mb_e5$TEAM_PITCHING_H+1.20501*mb_e5$TEAM_BATTING_SO+3.03598*mb_e5$TEAM_PITCHING_BB-8.40807*mb_e$TEAM_PITCHING_HR)

```

#####Building a linear regression and filling in NA's for DP
```{r mb_e5, eval=FALSE}

#build a regression model for DP's

DP <- lm(data=mb_e5, TEAM_FIELDING_DP~.)
summary(DP)

#eliminate wins, hitting, HBP

DP1 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_BASERUN_CS+TEAM_PITCHING_H+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP1)

#eliminate pitching hits allowed
DP2 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_BASERUN_CS+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP2)

#eliminate CS as it makes no sense
DP3 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP3)

#replace NA with values for DP
mb_e6 <- mb_e5
mb_e6$TEAM_FIELDING_DP[is.na(mb_e6$TEAM_FIELDING_DP)]<- round(158.8-0.1235*mb_e6$TEAM_BASERUN_SB +0.03320*mb_e6$TEAM_PITCHING_HR+0.02815*mb_e6$TEAM_PITCHING_BB - 0.006109*mb_e6$TEAM_PITCHING_SO - 0.06573*mb_e6$TEAM_FIELDING_E)

summary(mb_e6)
#only NA's left are HBP


```


## Data Imputation


```{r, eval=TRUE, include=TRUE}

# read EVALUATION data set
eval_data <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-evaluation-data.csv")

# read training data set
mb_e <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-training-data.csv")

#eliminate index column
# mb_e1 <- mb_e[,-1]
mb_e1 <- mb_e

#####Creating a new column for batting singles and eliminating hits for batting

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-3]
mb_e1 <- as.data.frame(mb_e1)

eval_data$TEAM_BATTING_1B <- as.numeric(eval_data$TEAM_BATTING_H - eval_data$TEAM_BATTING_2B - eval_data$TEAM_BATTING_3B - eval_data$TEAM_BATTING_HR)

# HITS is in second column in eval data
eval_data <- eval_data[,-2]

# ADD A DUMMY COLUMN TO EVAL DATA FOR TARGET WINS
eval_data$TARGET_WINS <- 0
```

#####Eliminate HBP, CS, and pitching HR's.

```{r, eval=TRUE, include=TRUE}
mb <- mb_e1[,-c(9,10,12)]
# summary(mb)

eval_data <- eval_data[,-c(8,9,11)]
# summary(eval_data)
```

#####Build model for batting SO using Gelman approach
```{r, eval=TRUE, include=TRUE}

#take out double plays + pitching SO + SB as data set is incomplete + Wins as they are not present in the evaluation data

BSO.1 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB -TARGET_WINS)
summary(BSO.1)

#eliminate doubles
BSO.2 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B)
summary(BSO.2)
vif(BSO.2)

# vif says remove TEAM_PITCHING_BB
BSO.3 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB)
summary(BSO.3)

# pvals say remove PITCHING_H
BSO.4 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(BSO.4)

vif(BSO.4)

##All p-values are low with a 686.8 F-statistic and adjusted R squared of 0.7236
#take a look
par(mfrow=c(2,2))
plot(BSO.2)

# ---------------------------------------
# function definition for impute function
impute <- function (a, a.impute){
  ifelse (is.na(a), a.impute,a)
}
# ---------------------------------------

#prediction function
pred.BSO <- round(predict(BSO.4, mb))
BSO.imp <- impute(mb$TEAM_BATTING_SO, pred.BSO)

# impute the evaluation data
pred_eval.BSO <- round(predict(BSO.4, eval_data))
eval.BSO.imp <- impute(eval_data$TEAM_BATTING_SO, pred_eval.BSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb$TEAM_BATTING_SO)
summary(BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb$TEAM_BATTING_SO, breaks = 200)
hist(BSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data$TEAM_BATTING_SO)
summary(eval.BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data$TEAM_BATTING_SO, breaks = 30)
hist(eval.BSO.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb1 <- mb
mb1$TEAM_BATTING_SO <- BSO.imp

eval_data.1 <- eval_data
eval_data.1$TEAM_BATTING_SO <- eval.BSO.imp

```


##### Build model for Pitching SO

```{r, eval=TRUE, include=TRUE}
#take out double plays + SB as data set is incomplete and wins as they are not present in evaluation data

PSO.1 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS)
summary(PSO.1)

vif(PSO.1)
# vif says remove TEAM_PITCHING_BB

PSO.2 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS - TEAM_PITCHING_BB)
summary(PSO.2)

vif(PSO.2)

#all low P value and F statistic of 4719 with adj R squared of 0.9952
#take a look
par(mfrow=c(2,2))
plot(PSO.2)

#place back in the data base with imputed data for SO's
pred.PSO <- round(predict(PSO.2, mb1))
PSO.imp <- impute(mb1$TEAM_PITCHING_SO, pred.PSO)

# impute the evaluation data
pred_eval.PSO <- round(predict(PSO.2, eval_data.1))
eval.PSO.imp <- impute(eval_data.1$TEAM_PITCHING_SO, pred_eval.PSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb1$TEAM_PITCHING_SO)
summary(PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb1$TEAM_PITCHING_SO, breaks = 200)
hist(PSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.1$TEAM_PITCHING_SO)
summary(eval.PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.1$TEAM_PITCHING_SO, breaks = 30)
hist(eval.PSO.imp, breaks = 30)

###################################################

# update dataframes with imputed values 

mb2 <- mb1
mb2$TEAM_PITCHING_SO <- PSO.imp

eval_data.2 <- eval_data.1
eval_data.2$TEAM_PITCHING_SO <- eval.PSO.imp
```

#####Build model for SB
```{r, eval=TRUE, include=TRUE}
#Take out DP as incomplete data and target wins
SB.1 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP - TARGET_WINS)
summary(SB.1)

#eliminate pitching BB's
SB.2 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS)
summary(SB.2)

#eliminate singles
SB.3 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS)
summary(SB.3)

#simplify the model by taking out pitching
SB.4 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.4)

#add singles back in
SB.5 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.5)

#eliminate doubles
SB.6 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B)
summary(SB.6)

#eliminate walks
SB.7 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B - TEAM_BATTING_BB)
summary(SB.7)

#all low P value and F statistic of 202.9 with adj R squared of 0.3427
#take a look
par(mfrow=c(2,2))
plot(SB.7)

#place back in the data base with imputed data for SB's
pred.SB <- round(predict(SB.7, mb2))
SB.imp <- impute(mb2$TEAM_BASERUN_SB, pred.SB)

# impute the evaluation data
pred_eval.SB <- round(predict(SB.7, eval_data.2))
eval.SB.imp <- impute(eval_data.2$TEAM_BASERUN_SB, pred_eval.SB)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb2$TEAM_BASERUN_SB)
summary(SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb2$TEAM_BASERUN_SB, breaks = 200)
hist(SB.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.2$TEAM_BASERUN_SB)
summary(eval.SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.2$TEAM_BASERUN_SB, breaks = 30)
hist(eval.SB.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb3 <- mb2
mb3$TEAM_BASERUN_SB <- SB.imp

eval_data.3 <- eval_data.2
eval_data.3$TEAM_BASERUN_SB <- eval.SB.imp
```

#####Build model to replace DP
```{r, eval=TRUE, include=TRUE}

#remove target wins
DP.1 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS)
summary(DP.1)

#remove batting 2B's
DP.2 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS - TEAM_BATTING_2B)
summary(DP.2)
# results show that EVERYTHING ELSE is statistically signficant, so:

# run vif to check for collinearity
vif(DP.2)
# results show TEAM_BATTING_SO should be removed

# remove TEAM_BATTING_SO
DP.3 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO)
summary(DP.3)
# p-value says remove TEAM_PITCHING_SO;


# remove TEAM_PITCHING_SO
DP.4 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TEAM_BATTING_2B -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO)
summary(DP.4)
vif(DP.4)
# P values and vif both indicate remove TEAM_PITCHING_BB

# remove TEAM_PITCHING_BB
DP.5 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(DP.5)
vif(DP.5)
# vif says remove TEAM_FIELDING_E; p-values all < .05 so remove TEAM_FIELDING_E

DP.6 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E)
summary(DP.6)
vif(DP.6)
# now no collinearity but p-values say remove TEAM_PITCHING_H

DP.7 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E - TEAM_PITCHING_H)
summary(DP.7)
vif(DP.7)
# no collinearity, all p-values < .05 so stop


#all low P value and F statistic of 255.8 with adj R squared of 0.3904
#take a look
par(mfrow=c(2,2))
plot(DP.7)


#place back in the data base with imputed data for SB's
# NOTE: Changed DP.4 to DP.7 here
pred.DP <- round(predict(DP.7, mb3))
DP.imp <- impute(mb3$TEAM_FIELDING_DP, pred.DP)

# impute the evaluation data
pred_eval.DP <- round(predict(DP.7, eval_data.3))
eval.DP.imp <- impute(eval_data.3$TEAM_FIELDING_DP, pred_eval.DP)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb3$TEAM_FIELDING_DP)
summary(DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb3$TEAM_FIELDING_DP, breaks = 200)
hist(DP.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.3$TEAM_FIELDING_DP)
summary(eval.DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.3$TEAM_FIELDING_DP, breaks = 30)
hist(eval.DP.imp, breaks = 30)
###################################################

# update data frames with imputed values
mb4 <- mb3
mb4$TEAM_FIELDING_DP <- DP.imp

eval_data.4 <- eval_data.3
eval_data.4$TEAM_FIELDING_DP <- eval.DP.imp
```


#####Eliminate unhistorical outliers - DO THIS FOR THE EVAL DATA AS WELL

```{r, eval=TRUE, include=TRUE}

# check rowcount before removal of outliers
nrow(mb4)
nrow(eval_data.4)

############## TEAM PITCHING_SO ############################
#most pitching SO's is 1450.  So delete all records with more than 1450 pitching SO's
mb5 <- mb4

# fixed error in this line: dataframe in 'which' call was mb1 so changed to mb5
mb5 <- mb5[which(mb5$TEAM_PITCHING_SO < 1450),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_SO < 1450),]

# check rowcount
nrow(mb5)
nrow(eval_data.4)

############ TEAM_PITCHING_H ##############################
#most ever hits by a team is 1730.  So delete all pitching hits >3000 to be conservative with the median
mb6 <- mb5
mb6 <- mb6[which(mb6$TEAM_PITCHING_H < 3001),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_H < 3001),]

# check rowcount
nrow(mb6)
nrow(eval_data.4)

############ TEAM_FIELDING_E ##############################
#most ever errors by a team is 639 by 1883 Philadelphia.  Prorating to 162 games gives a value of 1046.
mb7 <- mb6
mb7 <- mb7[which(mb7$TEAM_FIELDING_E < 1047),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_FIELDING_E < 1047),]

# ----------------------------------------------------------------------
# ----------------------------------------------------------------------

# check rowcount: result is 2172 => removed total of 104 rows
nrow(mb7)
nrow(eval_data.4)

dim(mb)-dim(mb7)

#we removed 104 rows total due to outliers in TRAINING data set.

# we removed 11 rows from the EVALUATION data set

# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(mb7) <- 1:nrow(mb7)
rownames(eval_data.4) <- 1:nrow(eval_data.4)

# drop INDEX column from training set
# mb7 <- mb7[,-1]

# now drop dummy column from evaluation data
# eval_data.4 <- eval_data.4[,-14]


# create CSV files containing updated data sets
#SMK commented out for now-doesn't work on a mac
#write.csv(mb7, file = "C:/SQLData/621-HW1-Clean-Data.csv", row.names = FALSE, col.names = TRUE)

#write.csv(eval_data.4, file = "C:/SQLData/621-HW1-Clean-EvalData-.csv", row.names = FALSE, col.names = TRUE)
```

# Part 3. Build Models


# Model 1: General Model Using Backward Selection


```{r part3}
library(car)

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```

```{r, eval = FALSE, echo = FALSE}

library(alr3)

# code for finding power transforms for skewed variables
hist(mb_clean$TEAM_PITCHING_BB, breaks = 200)

hist(mb_clean$TEAM_PITCHING_H, breaks = 200)

hist(mb_clean$TEAM_FIELDING_E, breaks = 200)

############### TEAM_PITCHING_BB ##############################

summary( powerTransform( cbind(TEAM_PITCHING_BB) ~ 1, mb_clean))

# TEAM_PITCHING_BB = log transform
PBB.T <- log(mb_clean$TEAM_PITCHING_BB)
hist(PBB.T, breaks = 200)

# ---------------------
############### PITCHING_H ###################################
summary( powerTransform( cbind(TEAM_PITCHING_H) ~ 1, mb_clean))
# TEAM_PITCHING_H = 1/y^3

# WORKS! 
PH.T <- 1/(mb_clean$TEAM_PITCHING_H ^ 3)
hist(PH.T, breaks = 200)


############## FIELDING_E #####################################
# -------------------------
summary( powerTransform( cbind(TEAM_FIELDING_E) ~ 1, mb_clean))
# TEAM_FIELDING_E = -1 = 1/y or try log

# WORKS!
FE.T <- 1/mb_clean$TEAM_FIELDING_E
hist(FE.T, breaks = 200)

# Now load transformed values into data set

mb.t <- mb_clean

mb.t$TEAM_PITCHING_BB <- PBB.T
mb.t$TEAM_PITCHING_H <- PH.T
mb.t$TEAM_FIELDING_E <- FE.T
```

## Build Models

### Model using all remaining variables as a starting point

__Yields r^2= 0.3347, Adj r^2 = 0.3322, F = 136__

```{r}

# keep the clean data set pure
mb <- mb_clean

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# p-value indicates remove TEAM_BATTING_1B

# --------------------
# remove TEAM_BATTING_1B
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_BATTING_1B)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_SO or PITCHING_SO, so remove PITCHING_SO per other models

# -------------------
#eliminate TEAM_PITCHING_SO
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO)
summary(model.3)

vif(model.3)

# vif says remove TEAM_BATTING_BB or PITCHINNG_BB so go with PITCHING_BB

# -------------------
#eliminate TEAM_PITCHING_BB
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# --------------------
# eliminate TEAM_BATTING_2B
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_BATTING_2B)
summary(model.5)

# p-values < .05 so check for collinearity
vif(model.5)

# no colinearity so STOP HERE
```


## DIAGNOSTICS

Plots are linear except for errors
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.5, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS
Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. Outliers at 1920, 1737, 393, 1515
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.5)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, SB, Pitch_H, Fielding_E, BATT_BB, BATT_SO, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.5)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.5$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```

Now cleanup data objects that are no longer required
```{r}
rm(model, model.2, model.3, model.4)
```



## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 1920, 1737, 393, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_clean[-c(1920, 1737, 393, 1515),]

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```


## Now refit first model from above: all variables

__Yields r^2= 0.3504, Adj r^2 = 0.348, F = 145.6__

```{r}

# keep the clean data set pure
mb <- mb_rem

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
vif(model)
# vif indicates remove TEAM_PITCHING_SO

# --------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# p-values say remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)
vif(model.4)

# pvals say remove TEAM_BATTING_2B
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_2B - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.5)
vif(model.5)


# vif and pvals OK so stop
```


## SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. __Might not be a good model__.
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.4)
```

## Plots show outliers so remove them and re-fit

Per Cooks Distance, remove items 1931, 391, 820, 1933, 835, 2124

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem2 <- mb[-c(1931, 391, 820, 1933, 835, 2124),]

# renumber rows
rownames(mb_rem2) <- 1:nrow(mb_rem2)
```

## Model using all remaining variables as a starting point

__Yields r^2= 0.3598, Adj r^2 = 0.3572, F = 134.4__

```{r}

# keep the clean data set pure
mb <- mb_rem2

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX )
summary(model)

# p-values all < .05 so check collinearity
vif(model)

# vif says remove TEAM_PITCHING_SO

# -------------------
#eliminate TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values OK so check collinearity
vif(model.2)
# vif says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_BATTING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# p-values says remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_BATTING_1B)
summary(model.4)

# p-values OK so check collinearity
vif(model.4)
# vif OK so STOP

# get MSE of residuals
anova(model.4)

```

## Diagnostics

Plots are linear except for errors
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. 
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, SB, Pitch_H, Fielding_E, PITCH_BB, PITCH_SO, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_2B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```


## Now try same model but with FIELD_E transformed by Box-Cox recommended power transform

```{r}
# TEAM_FIELDING_E: Box-cox yields -1 => 1/y

mb$TEAM_FIELDING_E <- 1/mb$TEAM_FIELDING_E
```


__Yields r^2= 0.3168, Adj r^2 = 0.3143, F = 124.8__

```{r}

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
# p-values all < .05 so check collinearity

vif(model)

# vif says remove TEAM_PITCHING_SO

# -------------------
#eliminate TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values says remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B)
summary(model.3)
vif(model.3)

# p-values says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_BATTING_BB
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B - TEAM_PITCHING_BB)
summary(model.4)

# p-values say remove doubles
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B - TEAM_PITCHING_BB - TEAM_BATTING_2B)
summary(model.5)

vif(model.5)

# vif OK so STOP

# turn off scientific formatting of results
options(scipen=999)
model.5
anova(model.5)
```

## Diagnostics

Plots are all linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.5, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in lower left end but not drastic; Residuals not all within 2 std devs.  
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.5)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, BB, SO, SB, Pitch_H, Fielding_E
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.5)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be slightly skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.5$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```


```{r}
# clean up objects in memory
rm(list = ls())
```



# Model 2: Total Bases






```{r}
library(car)

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```


## Build a model with Total Bases added and all of the other hitting vars removed

First, create the new variable and discard its components

```{r}
mb_t <- mb_clean

mb_t$TOTAL_BASES <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR)

# plot histogram to check shape of distribution
par(mfrow = c(1,1))
hist(mb_t$TOTAL_BASES, breaks = 200)

# now drop 1B, 2B, 3B, HR
mb_tb <- mb_t[,c(1, 2, 6, 7, 8, 9, 10, 11, 12, 13, 15)]

###################################################################
# check correlation with WINS and run simple linear model
cor(mb_tb$TARGET_WINS, mb_tb$TOTAL_BASES) 
mtest <- lm(data=mb_tb, TARGET_WINS ~ TOTAL_BASES)
summary(mtest) 

# shows .381 correlation and Adj R^2 of 0.1453

plot(mb_tb$TARGET_WINS ~ mb_tb$TOTAL_BASES)
abline(lm(mb_tb$TARGET_WINS ~ mb_tb$TOTAL_BASES), lty=2)
# plot doesn't show unusual relationship
######################################################################
```

__Yields r^2= 0.3175, Adj. R^2 = 0.3153, F = 143.8__

```{r}
# fit model
model <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX)
summary(model)

# All p-values < .05 so check collinearity
vif(model)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO. CHoose PITCHING_SO as in other models

# ----------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.3 <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# All p-values < .05 so check collinearity
vif(model.3)
# no further collinearity issues so STOP

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E  shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots look good except for outliers. Lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals except for outliers, most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.3)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for BASERUN_SB, PITCHING_H, PITCHING_SO, FIELDING_E, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.3)
par(mfrow=c(2,2))

plot(mb_tb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")

plot(mb_tb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Might be some skew due to outliers
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.3$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tb$TARGET_WINS),lty=2)
```


## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 1920, 1737, 393, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_tb[-c(1920, 1737, 393, 1515),]

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```

## Now refit first model from above: all variables


__Yields r^2= 0.3287, Adj r^2 = 0.3265, F = 151.1__

```{r}

# keep the clean data set pure
mb <- mb_rem

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# pvals all < .05 so check collinearity

vif(model)
# vif indicates remove TEAM_BATTING_SO and PITCHING_SO. CHoose PITCHING_SO again

# --------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_BB

# -------------------
#eliminate TEAM_BATTING_BB or PITCHING_BB so choose PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_SO)
summary(model.3)

# p-values OK so check collinearity
vif(model.3)
# vif and pvals OK so STOP HERE

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Some residuals appear to be outside 2 std devs. __Might not be a good model__. Outliers at 1528, 1922, 820, 1933, 1733, 835
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.3)
```


### Plots show outliers so remove them and re-fit

Per Cooks Distance, remove 1528, 1922, 820, 1933, 1733, 835

```{r}
############ SECOND SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem2 <- mb[-c(1528, 1922, 820, 1933, 1733, 835),]

# renumber rows
rownames(mb_rem2) <- 1:nrow(mb_rem2)
```

### Model using all remaining variables as a starting point

__Yields r^2= 0.33655, Adj r^2 = 0.3343, F = 156__

```{r}

# keep the clean data set pure
mb <- mb_rem2

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
# p-values all < .05 so check collinearity

vif(model)

# vif says remove TEAM_BATTING_SO

# -------------------
#eliminate TEAM_BATTING_SO or PITCHING_SO so choose PITCHING_SO again
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_BB or PITCHING_BB so choose PITCHING_BB again

# -------------------
#eliminate TEAM_PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

vif(model.3)
# p-values and VIF OK so STOP

anova(model.3)

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted but only for extreme outliers. Normal QQ looks very good; Some Residuals appear to be outside 2 std devs
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.3)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for BASERUN_SB, PITCHING_H, PITCHING_SO, FIELDING_E, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.3)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Might be some skew due to outliers
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.3$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```
# -------------------------------------------------------------------------------------------
# -------------------------------------------------------------------------------------------

## Now try same model but with FIELD_E transformed according to Box-Cox

```{r}
# TEAM_FIELDING_E: Box-Cox yields power xform of -1 => 1/y

mb$TEAM_FIELDING_E <- 1/mb$TEAM_FIELDING_E
```

Now refit first model from above: all variables

## Model using all remaining variables as a starting point

__Yields r^2= 0.3048, Adj r^2 = 0.3029, F = 157.5__

```{r}

# fit model
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

vif(model.2)
# vif says remove TEAM_BATTING_SO or PITCHING_SO so discard PITCHING_SO again

# ----------------------
# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

vif(model.3)
# vif say remove TEAM_BATTING_BB or PITCHING_BB - go with PITCHING_BB again

# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

options(scipen=999)
model.4
anova(model.4)

# no collinearity so STOP
# check 95% confidence intervals for coefficients
confint(model.4)

```

## Diagnostics

Plots show all variables are linear to response
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at both ends; normal distribution of residuals; many residuals outside of 2 std devs __PROBABLY NOT A GOOD MODEL__
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```

```{r}
# clean up objects in memory
rm(list = ls())
```






# Model 3: Total Bases PLUS

```{r}
library(car)

# read clean data set from Github

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```

## Build a model with Total Bases + SB + BB added and all of the other hitting vars removed

```{r}

# create new variable and drop its components
mb_t <- mb_clean

mb_t$TB_PLUS <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR) + 
                       mb_clean$TEAM_BATTING_BB + mb_clean$TEAM_BASERUN_SB

par(mfrow = c(1,1))
hist(mb_t$TB_PLUS, breaks = 200)

# now drop 1B, 2B, 3B, HR, BB, SB
mb_tbp <- mb_t[,c(1, 2, 7, 9, 10, 11, 12, 13, 15)]

###################################################################
# check correlation with WINS and run simple linear model
cor(mb_tbp$TARGET_WINS, mb_tbp$TB_PLUS) 
mtest <- lm(data=mb_tbp, TARGET_WINS ~ TB_PLUS)
summary(mtest) 

# shows .448 correlation and Adj R^2 of 0.2002 => better than TOTAL_BASES

plot(mb_tbp$TARGET_WINS ~ mb_tbp$TB_PLUS)
abline(lm(mb_tbp$TARGET_WINS ~ mb_tbp$TB_PLUS), lty=2)
# plot doesn't show unusual relationship
######################################################################
```



__Yields r^2= 0.2845, Adj. R^2 = 0.2832, F = 215.4__

```{r}
# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO, so try removing PITCHING_SO

# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

# p-vals say remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

# no collinearity so STOP

```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: outliers at 2012, 1820, 859. Lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals except for outliers, most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```




## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 836, 821, 1920, 1737, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_tbp[-c(836, 821, 1920, 1737, 1515),]

# save first data set
mb_tbp_orig <- mb_tbp

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```

## Now refit first model from above: all variables

__Yields r^2= 0.2944, Adj r^2 = 0.2931, F = 225.5__

```{r}

# keep the clean data set pure
mb_tbp <- mb_rem

# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ .- INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO, so remove PITCHING_SO again

# ----------------------
# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

# p-vals say remove TEAM_PITCHING_BB


# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

anova(model)

# no collinearity so STOP

```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals; most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```




## Now try same model but with FIELD_E transformed using Box-Cox

```{r}
# TEAM_FIELDING_E: Box-cox says -1 power transform => 1/y 

mb_tbp$TEAM_FIELDING_E <- 1/mb_tbp$TEAM_FIELDING_E
```


Now refit first model from above: Start with all variables

__Yields r^2= 0.2932, Adj r^2 = 0.2919, F = 224.3__

```{r}

# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB)
summary(model.2)

# p-values say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(model.3)

# p-vals say remove TEAM_PITCHING_SO

# ----------------------
# remove TEAM_PITCHING_SO
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

# no collinearity so STOP
options(scipen=999)
model.4
anova(model.4)
```

## Diagnostics

Plots show all variables are linear to response
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals; most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_PITCHING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```

```{r}
# clean up objects in memory
rm(list = ls())
```


# Model 4: Sabermetrics Model


```{r model4}
library(car)
```

```{r}

# read EVALUATION data set
eval_data <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-evaluation-data.csv")

# read training data set
mb_e <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-training-data.csv")

#eliminate index column
# mb_e1 <- mb_e[,-1]
mb_e1 <- mb_e

#####Creating a new column for batting singles and eliminating hits for batting

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-3]
mb_e1 <- as.data.frame(mb_e1)

eval_data$TEAM_BATTING_1B <- as.numeric(eval_data$TEAM_BATTING_H - eval_data$TEAM_BATTING_2B - eval_data$TEAM_BATTING_3B - eval_data$TEAM_BATTING_HR)

# HITS is in second column in eval data
eval_data <- eval_data[,-2]

# ADD A DUMMY COLUMN TO EVAL DATA FOR TARGET WINS
eval_data$TARGET_WINS <- 0
```

#####Eliminate HBP, CS, and pitching HR's.

```{r}
mb <- mb_e1[,-c(9,10,12)]
# summary(mb)

eval_data <- eval_data[,-c(8,9,11)]
# summary(eval_data)
```


#####Build model for batting SO using Gelman approach
```{r}

#take out double plays + pitching SO + SB as data set is incomplete + Wins as they are not present in the evaluation data

BSO.1 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB -TARGET_WINS)
summary(BSO.1)

#eliminate doubles
BSO.2 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B)
summary(BSO.2)
vif(BSO.2)

# vif says remove TEAM_PITCHING_BB
BSO.3 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB)
summary(BSO.3)

# pvals say remove PITCHING_H
BSO.4 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(BSO.4)

vif(BSO.4)

##All p-values are low with a 686.8 F-statistic and adjusted R squared of 0.7236
#take a look
par(mfrow=c(2,2))
plot(BSO.2)

# ---------------------------------------
# function definition for impute function
impute <- function (a, a.impute){
  ifelse (is.na(a), a.impute,a)
}
# ---------------------------------------

#prediction function
pred.BSO <- round(predict(BSO.4, mb))
BSO.imp <- impute(mb$TEAM_BATTING_SO, pred.BSO)

# impute the evaluation data
pred_eval.BSO <- round(predict(BSO.4, eval_data))
eval.BSO.imp <- impute(eval_data$TEAM_BATTING_SO, pred_eval.BSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb$TEAM_BATTING_SO)
summary(BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb$TEAM_BATTING_SO, breaks = 200)
hist(BSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data$TEAM_BATTING_SO)
summary(eval.BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data$TEAM_BATTING_SO, breaks = 30)
hist(eval.BSO.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb1 <- mb
mb1$TEAM_BATTING_SO <- BSO.imp

eval_data.1 <- eval_data
eval_data.1$TEAM_BATTING_SO <- eval.BSO.imp

```


### Build model for Pitching SO

```{r}
#take out double plays + SB as data set is incomplete and wins as they are not present in evaluation data

PSO.1 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS)
summary(PSO.1)

vif(PSO.1)
# vif says remove TEAM_PITCHING_BB

PSO.2 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS - TEAM_PITCHING_BB)
summary(PSO.2)

vif(PSO.2)

#all low P value and F statistic of 4719 with adj R squared of 0.9952
#take a look
par(mfrow=c(2,2))
plot(PSO.2)

#place back in the data base with imputed data for SO's
pred.PSO <- round(predict(PSO.2, mb1))
PSO.imp <- impute(mb1$TEAM_PITCHING_SO, pred.PSO)

# impute the evaluation data
pred_eval.PSO <- round(predict(PSO.2, eval_data.1))
eval.PSO.imp <- impute(eval_data.1$TEAM_PITCHING_SO, pred_eval.PSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb1$TEAM_PITCHING_SO)
summary(PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb1$TEAM_PITCHING_SO, breaks = 200)
hist(PSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.1$TEAM_PITCHING_SO)
summary(eval.PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.1$TEAM_PITCHING_SO, breaks = 30)
hist(eval.PSO.imp, breaks = 30)

###################################################

# update dataframes with imputed values 

mb2 <- mb1
mb2$TEAM_PITCHING_SO <- PSO.imp

eval_data.2 <- eval_data.1
eval_data.2$TEAM_PITCHING_SO <- eval.PSO.imp
```

#####Build model for SB
```{r}
#Take out DP as incomplete data and target wins
SB.1 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP - TARGET_WINS)
summary(SB.1)

#eliminate pitching BB's
SB.2 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS)
summary(SB.2)

#eliminate singles
SB.3 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS)
summary(SB.3)

#simplify the model by taking out pitching
SB.4 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.4)

#add singles back in
SB.5 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.5)

#eliminate doubles
SB.6 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B)
summary(SB.6)

#eliminate walks
SB.7 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B - TEAM_BATTING_BB)
summary(SB.7)

#all low P value and F statistic of 202.9 with adj R squared of 0.3427
#take a look
par(mfrow=c(2,2))
plot(SB.7)

#place back in the data base with imputed data for SB's
pred.SB <- round(predict(SB.7, mb2))
SB.imp <- impute(mb2$TEAM_BASERUN_SB, pred.SB)

# impute the evaluation data
pred_eval.SB <- round(predict(SB.7, eval_data.2))
eval.SB.imp <- impute(eval_data.2$TEAM_BASERUN_SB, pred_eval.SB)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb2$TEAM_BASERUN_SB)
summary(SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb2$TEAM_BASERUN_SB, breaks = 200)
hist(SB.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.2$TEAM_BASERUN_SB)
summary(eval.SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.2$TEAM_BASERUN_SB, breaks = 30)
hist(eval.SB.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb3 <- mb2
mb3$TEAM_BASERUN_SB <- SB.imp

eval_data.3 <- eval_data.2
eval_data.3$TEAM_BASERUN_SB <- eval.SB.imp
```

#####Build model to replace DP
```{r}

#remove target wins
DP.1 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS)
summary(DP.1)

#remove batting 2B's
DP.2 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS - TEAM_BATTING_2B)
summary(DP.2)
# results show that EVERYTHING ELSE is statistically signficant, so:

# run vif to check for collinearity
vif(DP.2)
# results show TEAM_BATTING_SO should be removed

# remove TEAM_BATTING_SO
DP.3 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO)
summary(DP.3)
# p-value says remove TEAM_PITCHING_SO;


# remove TEAM_PITCHING_SO
DP.4 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TEAM_BATTING_2B -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO)
summary(DP.4)
vif(DP.4)
# P values and vif both indicate remove TEAM_PITCHING_BB

# remove TEAM_PITCHING_BB
DP.5 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(DP.5)
vif(DP.5)
# vif says remove TEAM_FIELDING_E; p-values all < .05 so remove TEAM_FIELDING_E

DP.6 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E)
summary(DP.6)
vif(DP.6)
# now no collinearity but p-values say remove TEAM_PITCHING_H

DP.7 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E - TEAM_PITCHING_H)
summary(DP.7)
vif(DP.7)
# no collinearity, all p-values < .05 so stop


#all low P value and F statistic of 255.8 with adj R squared of 0.3904
#take a look
par(mfrow=c(2,2))
plot(DP.7)


#place back in the data base with imputed data for SB's
# NOTE: Changed DP.4 to DP.7 here
pred.DP <- round(predict(DP.7, mb3))
DP.imp <- impute(mb3$TEAM_FIELDING_DP, pred.DP)

# impute the evaluation data
pred_eval.DP <- round(predict(DP.7, eval_data.3))
eval.DP.imp <- impute(eval_data.3$TEAM_FIELDING_DP, pred_eval.DP)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb3$TEAM_FIELDING_DP)
summary(DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb3$TEAM_FIELDING_DP, breaks = 200)
hist(DP.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.3$TEAM_FIELDING_DP)
summary(eval.DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.3$TEAM_FIELDING_DP, breaks = 30)
hist(eval.DP.imp, breaks = 30)
###################################################

# update data frames with imputed values
mb4 <- mb3
mb4$TEAM_FIELDING_DP <- DP.imp

eval_data.4 <- eval_data.3
eval_data.4$TEAM_FIELDING_DP <- eval.DP.imp
```


#####Eliminate unhistorical outliers - DO THIS FOR THE EVAL DATA AS WELL

```{r}

# check rowcount before removal of outliers
nrow(mb4)
nrow(eval_data.4)

```

```{r}


############## TEAM PITCHING_SO ############################
#most pitching SO's is 1450.  So delete all records with more than 1450 pitching SO's
mb5 <- mb4

# fixed error in this line: dataframe in 'which' call was mb1 so changed to mb5
mb5 <- mb5[which(mb5$TEAM_PITCHING_SO < 1450),]

# eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_SO < 1450),]

# check rowcount
nrow(mb5)
nrow(eval_data.4)
```

```{r}



############ TEAM_PITCHING_H ##############################
#most ever hits by a team is 1730.  So delete all pitching hits >3000 to be conservative with the median
mb6 <- mb5
mb6 <- mb6[which(mb6$TEAM_PITCHING_H < 3001),]

# eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_H < 3001),]

# check rowcount
nrow(mb6)
nrow(eval_data.4)

```

```{r}

############ TEAM_FIELDING_E ##############################
#most ever errors by a team is 639 by 1883 Philadelphia.  Prorating to 162 games gives a value of 1046.
mb7 <- mb6
mb7 <- mb7[which(mb7$TEAM_FIELDING_E < 1047),]

# eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_FIELDING_E < 1047),]

# ----------------------------------------------------------------------
# ----------------------------------------------------------------------

# check rowcount: result is 2172 => removed total of 104 rows
nrow(mb7)
nrow(eval_data.4)

dim(mb)-dim(mb7)

#we removed 104 rows total due to outliers in TRAINING data set.

# we removed 11 rows from the EVALUATION data set

# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(mb7) <- 1:nrow(mb7)
rownames(eval_data.4) <- 1:nrow(eval_data.4)

# drop INDEX column from training set
# mb7 <- mb7[,-1]

# now drop dummy column from evaluation data
# eval_data.4 <- eval_data.4[,-14]

# create CSV files containing updated data sets

```

```{r writecsv1, eval=FALSE}

write.csv(mb7, file = "C:/SQLData/621-HW1-Clean-Data.csv", row.names = FALSE, col.names = TRUE)

write.csv(eval_data.4, file = "C:/SQLData/621-HW1-Clean-EvalData-.csv", row.names = FALSE, col.names = TRUE)
```


# Model 5: Box-Cox First

```{r model5 load-libaries, eval=TRUE, include=FALSE}
library(car)
library(fBasics)
library(knitr)
library(car)
library(corrplot)
library(alr3)
library(MASS)
```


```{r,eval=TRUE, include=FALSE}

# read EVALUATION data set
eval_data <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-evaluation-data.csv")

# read training data set
mb_e <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-training-data.csv")

#eliminate index column
# mb_e1 <- mb_e[,-1]
mb_e1 <- mb_e

#####Creating a new column for batting singles and eliminating hits for batting

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-3]
mb_e1 <- as.data.frame(mb_e1)

eval_data$TEAM_BATTING_1B <- as.numeric(eval_data$TEAM_BATTING_H - eval_data$TEAM_BATTING_2B - eval_data$TEAM_BATTING_3B - eval_data$TEAM_BATTING_HR)

# HITS is in second column in eval data
eval_data <- eval_data[,-2]

# ADD A DUMMY COLUMN TO EVAL DATA FOR TARGET WINS
eval_data$TARGET_WINS <- 0
```

##### Eliminate HBP, CS, and pitching HR's.

```{r,eval=TRUE,include=FALSE}
mb <- mb_e1[,-c(9,10,12)]
# summary(mb)

eval_data <- eval_data[,-c(8,9,11)]
# summary(eval_data)
```


##### Build model for batting SO using Gelman approach
```{r,eval=TRUE,include=FALSE}

#take out double plays + pitching SO + SB as data set is incomplete + Wins as they are not present in the evaluation data

BSO.1 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB -TARGET_WINS)
summary(BSO.1)

#eliminate doubles
BSO.2 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B)
summary(BSO.2)
vif(BSO.2)

# vif says remove TEAM_PITCHING_BB
BSO.3 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB)
summary(BSO.3)

# pvals say remove PITCHING_H
BSO.4 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(BSO.4)

vif(BSO.4)

##All p-values are low with a 686.8 F-statistic and adjusted R squared of 0.7236
#take a look
par(mfrow=c(2,2))
plot(BSO.2)

# ---------------------------------------
# function definition for impute function
impute <- function (a, a.impute){
  ifelse (is.na(a), a.impute,a)
}
# ---------------------------------------

#prediction function
pred.BSO <- round(predict(BSO.4, mb))
BSO.imp <- impute(mb$TEAM_BATTING_SO, pred.BSO)

# impute the evaluation data
pred_eval.BSO <- round(predict(BSO.4, eval_data))
eval.BSO.imp <- impute(eval_data$TEAM_BATTING_SO, pred_eval.BSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb$TEAM_BATTING_SO)
summary(BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb$TEAM_BATTING_SO, breaks = 200)
hist(BSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data$TEAM_BATTING_SO)
summary(eval.BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data$TEAM_BATTING_SO, breaks = 30)
hist(eval.BSO.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb1 <- mb
mb1$TEAM_BATTING_SO <- BSO.imp

eval_data.1 <- eval_data
eval_data.1$TEAM_BATTING_SO <- eval.BSO.imp

```


### Build model for Pitching SO

```{r,eval=TRUE, include=FALSE}
#take out double plays + SB as data set is incomplete and wins as they are not present in evaluation data

PSO.1 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS)
summary(PSO.1)

vif(PSO.1)
# vif says remove TEAM_PITCHING_BB

PSO.2 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS - TEAM_PITCHING_BB)
summary(PSO.2)

vif(PSO.2)

#all low P value and F statistic of 4719 with adj R squared of 0.9952
#take a look
par(mfrow=c(2,2))
plot(PSO.2)

#place back in the data base with imputed data for SO's
pred.PSO <- round(predict(PSO.2, mb1))
PSO.imp <- impute(mb1$TEAM_PITCHING_SO, pred.PSO)

# impute the evaluation data
pred_eval.PSO <- round(predict(PSO.2, eval_data.1))
eval.PSO.imp <- impute(eval_data.1$TEAM_PITCHING_SO, pred_eval.PSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb1$TEAM_PITCHING_SO)
summary(PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb1$TEAM_PITCHING_SO, breaks = 200)
hist(PSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.1$TEAM_PITCHING_SO)
summary(eval.PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.1$TEAM_PITCHING_SO, breaks = 30)
hist(eval.PSO.imp, breaks = 30)

###################################################

# update dataframes with imputed values 

mb2 <- mb1
mb2$TEAM_PITCHING_SO <- PSO.imp

eval_data.2 <- eval_data.1
eval_data.2$TEAM_PITCHING_SO <- eval.PSO.imp
```

#####Build model for SB
```{r,eval=TRUE,include=FALSE}
#Take out DP as incomplete data and target wins
SB.1 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP - TARGET_WINS)
summary(SB.1)

#eliminate pitching BB's
SB.2 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS)
summary(SB.2)

#eliminate singles
SB.3 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS)
summary(SB.3)

#simplify the model by taking out pitching
SB.4 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.4)

#add singles back in
SB.5 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.5)

#eliminate doubles
SB.6 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B)
summary(SB.6)

#eliminate walks
SB.7 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B - TEAM_BATTING_BB)
summary(SB.7)

#all low P value and F statistic of 202.9 with adj R squared of 0.3427
#take a look
par(mfrow=c(2,2))
plot(SB.7)

#place back in the data base with imputed data for SB's
pred.SB <- round(predict(SB.7, mb2))
SB.imp <- impute(mb2$TEAM_BASERUN_SB, pred.SB)

# impute the evaluation data
pred_eval.SB <- round(predict(SB.7, eval_data.2))
eval.SB.imp <- impute(eval_data.2$TEAM_BASERUN_SB, pred_eval.SB)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb2$TEAM_BASERUN_SB)
summary(SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb2$TEAM_BASERUN_SB, breaks = 200)
hist(SB.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.2$TEAM_BASERUN_SB)
summary(eval.SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.2$TEAM_BASERUN_SB, breaks = 30)
hist(eval.SB.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb3 <- mb2
mb3$TEAM_BASERUN_SB <- SB.imp

eval_data.3 <- eval_data.2
eval_data.3$TEAM_BASERUN_SB <- eval.SB.imp
```

#####Build model to replace DP
```{r,eval=TRUE,include=FALSE}

#remove target wins
DP.1 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS)
summary(DP.1)

#remove batting 2B's
DP.2 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS - TEAM_BATTING_2B)
summary(DP.2)
# results show that EVERYTHING ELSE is statistically signficant, so:

# run vif to check for collinearity
vif(DP.2)
# results show TEAM_BATTING_SO should be removed

# remove TEAM_BATTING_SO
DP.3 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO)
summary(DP.3)
# p-value says remove TEAM_PITCHING_SO;


# remove TEAM_PITCHING_SO
DP.4 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TEAM_BATTING_2B -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO)
summary(DP.4)
vif(DP.4)
# P values and vif both indicate remove TEAM_PITCHING_BB

# remove TEAM_PITCHING_BB
DP.5 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(DP.5)
vif(DP.5)
# vif says remove TEAM_FIELDING_E; p-values all < .05 so remove TEAM_FIELDING_E

DP.6 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E)
summary(DP.6)
vif(DP.6)
# now no collinearity but p-values say remove TEAM_PITCHING_H

DP.7 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E - TEAM_PITCHING_H)
summary(DP.7)
vif(DP.7)
# no collinearity, all p-values < .05 so stop


#all low P value and F statistic of 255.8 with adj R squared of 0.3904
#take a look
par(mfrow=c(2,2))
plot(DP.7)


#place back in the data base with imputed data for SB's
# NOTE: Changed DP.4 to DP.7 here
pred.DP <- round(predict(DP.7, mb3))
DP.imp <- impute(mb3$TEAM_FIELDING_DP, pred.DP)

# impute the evaluation data
pred_eval.DP <- round(predict(DP.7, eval_data.3))
eval.DP.imp <- impute(eval_data.3$TEAM_FIELDING_DP, pred_eval.DP)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb3$TEAM_FIELDING_DP)
summary(DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb3$TEAM_FIELDING_DP, breaks = 200)
hist(DP.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.3$TEAM_FIELDING_DP)
summary(eval.DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.3$TEAM_FIELDING_DP, breaks = 30)
hist(eval.DP.imp, breaks = 30)
###################################################

# update data frames with imputed values
mb4 <- mb3
mb4$TEAM_FIELDING_DP <- DP.imp

eval_data.4 <- eval_data.3
eval_data.4$TEAM_FIELDING_DP <- eval.DP.imp
```


#####Eliminate unhistorical outliers - DO THIS FOR THE EVAL DATA AS WELL

```{r,eval=TRUE,include=FALSE}

# check rowcount before removal of outliers
nrow(mb4)
nrow(eval_data.4)

############## TEAM PITCHING_SO ############################
#most pitching SO's is 1450.  So delete all records with more than 1450 pitching SO's
mb5 <- mb4

# fixed error in this line: dataframe in 'which' call was mb1 so changed to mb5
mb5 <- mb5[which(mb5$TEAM_PITCHING_SO < 1450),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_SO < 1450),]

# check rowcount
nrow(mb5)
nrow(eval_data.4)

############ TEAM_PITCHING_H ##############################
#most ever hits by a team is 1730.  So delete all pitching hits >3000 to be conservative with the median
mb6 <- mb5
mb6 <- mb6[which(mb6$TEAM_PITCHING_H < 3001),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_H < 3001),]

# check rowcount
nrow(mb6)
nrow(eval_data.4)

############ TEAM_FIELDING_E ##############################
#most ever errors by a team is 639 by 1883 Philadelphia.  Prorating to 162 games gives a value of 1046.
mb7 <- mb6
mb7 <- mb7[which(mb7$TEAM_FIELDING_E < 1047),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_FIELDING_E < 1047),]

# ----------------------------------------------------------------------
# ----------------------------------------------------------------------

# check rowcount: result is 2172 => removed total of 104 rows
nrow(mb7)
nrow(eval_data.4)

dim(mb)-dim(mb7)

#we removed 104 rows total due to outliers in TRAINING data set.

# we removed 11 rows from the EVALUATION data set

# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(mb7) <- 1:nrow(mb7)
rownames(eval_data.4) <- 1:nrow(eval_data.4)

# drop INDEX column from training set
# mb7 <- mb7[,-1]

# now drop dummy column from evaluation data
# eval_data.4 <- eval_data.4[,-14]

```

```{r writecsv2, eval=FALSE}
# create CSV files containing updated data sets
write.csv(mb7, file = "/Users/scottkarr/IS621Summer2016/HW1/621-HW1-Clean-Data.csv", row.names = FALSE, col.names = TRUE)

write.csv(eval_data.4, file = "/Users/scottkarr/IS621Summer2016/HW1/621-HW1-Clean-EvalData-.csv", row.names = FALSE, col.names = TRUE)
```

### SINGLE PREDICTOR ANALYSIS & TRANSFORMATIONS
#####Model SMK Generlized Equation
Review descriptive statistics to confirm each variable is within acceptable bounds and 
contains no missing data.  Review Density plots of 13 variables for skewness to identify
which may require transformation.  

```{r full, eval = TRUE,include=FALSE, echo=FALSE}
#assign model to "clean" data set 
lm.smk <- mb7
#remove bad leverage points from diagnostic tests
lm.smk <- lm.smk[-c(1737,1920,226,391,385,1702,840,602,1928,1937,2109,2128,269,711,125),]
# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(lm.smk) <- 1:nrow(lm.smk)
nrow(lm.smk)

#SINGLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.1B <- lm.smk$TEAM_BATTING_1B
m1 <- lm(mW~mb.1B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.1B),main="Singles");rug(mb.1B)
plot(mb.1B,StanRes1,xlab="Singles",ylab="Standardized Residuals");abline(lsfit(mb.1B,StanRes1),lty=2,col=2)
qqnorm(mb.1B,ylab="Y");qqline(mb.1B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.1B, family="bcPower")$lambda# round(-1.983968) => -2 => 1/(y^2)
tmb.1B <- mb.1B^(-2)
plot(density(tmb.1B),main="Singles");rug(tmb.1B)
plot(tmb.1B,StanRes1,xlab="Singles",ylab="Standardized Residuals");abline(lsfit(tmb.1B,StanRes1),lty=2,col=2)
qqnorm(tmb.1B,ylab="Y");qqline(tmb.1B,lty=2,col=2)

#DOUBLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.2B <- lm.smk$TEAM_BATTING_2B
m1 <- lm(mW~mb.2B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.2B),main="Doubles");rug(mb.2B)
plot(mb.2B,StanRes1,xlab="Doubles",ylab="Standardized Residuals");abline(lsfit(mb.2B,StanRes1),lty=2,col=2)
qqnorm(mb.2B,ylab="Y");qqline(mb.2B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.2B, family="bcPower")$lambda# round(0.5362315) => y^.5 => sqrt(y)
tmb.2B <- sqrt(mb.2B)
plot(density(tmb.2B),main="Doubles");rug(tmb.2B)
plot(tmb.2B,StanRes1,xlab="Doubles",ylab="Standardized Residuals");abline(lsfit(tmb.2B,StanRes1),lty=2,col=2)
qqnorm(tmb.2B,ylab="Y");qqline(tmb.2B,lty=2,col=2)

#TRIPLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.3B <- lm.smk$TEAM_BATTING_3B
m1 <- lm(mW~mb.3B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.3B),main="Triples");rug(mb.3B)
plot(mb.3B,StanRes1,xlab="Triples",ylab="Standardized Residuals");abline(lsfit(mb.3B,StanRes1),lty=2,col=2)
qqnorm(mb.3B,ylab="Y");qqline(mb.3B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.3B, family="bcPower")$lambda# round(-0.03308475) => (1/y^30)
tmb.3B <- mb.3B^(-30)
plot(density(tmb.3B),main="Triples");rug(tmb.3B)
plot(tmb.3B,StanRes1,xlab="Triples",ylab="Standardized Residuals");abline(lsfit(tmb.3B,StanRes1),lty=2,col=2)
qqnorm(tmb.3B,ylab="Y");qqline(tmb.3B,lty=2,col=2)

#HOMERUNS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.HR <- lm.smk$TEAM_BATTING_HR
m1 <- lm(mW~mb.HR, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.HR),main="Homeruns");rug(mb.HR)
plot(mb.HR,StanRes1,xlab="Homeruns",ylab="Standardized Residuals");abline(lsfit(mb.HR,StanRes1),lty=2,col=2)
qqnorm(mb.HR,ylab="Y");qqline(mb.HR,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.HR, family="bcPower")$lambda# round(0.6348318) => y^(2/3)
tmb.HR <- mb.HR^(2/3)
plot(density(tmb.HR),main="Homeruns");rug(tmb.HR)
plot(tmb.HR,StanRes1,xlab="Homeruns",ylab="Standardized Residuals");abline(lsfit(tmb.HR,StanRes1),lty=2,col=2)
qqnorm(tmb.HR,ylab="Y");qqline(tmb.HR,lty=2,col=2)

#SLUGGING
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SL <- lm.smk$TEAM_BATTING_HR + 2 * lm.smk$TEAM_BATTING_3B 
m1 <- lm(mW~mb.SL, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SL),main="Slugging");rug(mb.SL)
plot(mb.SL,StanRes1,xlab="Slugging",ylab="Standardized Residuals");abline(lsfit(mb.SL,StanRes1),lty=2,col=2)
qqnorm(mb.SL,ylab="Y");qqline(mb.SL,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SL, family="bcPower")$lambda# round(0.5562253) => y^(.5)
tmb.SL <- sqrt(mb.SL)
plot(density(tmb.SL),main="Slugging");rug(tmb.SL)
plot(tmb.SL,StanRes1,xlab="Slugging",ylab="Standardized Residuals");abline(lsfit(tmb.SL,StanRes1),lty=2,col=2)
qqnorm(tmb.SL,ylab="Y");qqline(tmb.SL,lty=2,col=2)

#WALKS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.BB <- lm.smk$TEAM_BATTING_BB
m1 <- lm(mW~mb.BB, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.BB),main="Walks");rug(mb.BB)
plot(mb.BB,StanRes1,xlab="Walks",ylab="Standardized Residuals");abline(lsfit(mb.BB,StanRes1),lty=2,col=2)
qqnorm(mb.BB,ylab="Y");qqline(mb.BB,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.BB, family="bcPower")$lambda# round(1.434735) => y^(3/2)
tmb.BB <- mb.BB^(3/2)
plot(density(tmb.BB),main="Walks");rug(tmb.BB)
plot(tmb.BB,StanRes1,xlab="Walks",ylab="Standardized Residuals");abline(lsfit(tmb.BB,StanRes1),lty=2,col=2)
qqnorm(tmb.BB,ylab="Y");qqline(tmb.BB,lty=2,col=2)

#STRIKEOUTS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SO <- lm.smk$TEAM_BATTING_SO
m1 <- lm(mW~mb.SO, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SO),main="Strikeouts");rug(mb.SO)
plot(mb.SO,StanRes1,xlab="Strikeouts",ylab="Standardized Residuals");abline(lsfit(mb.SO,StanRes1),lty=2,col=2)
qqnorm(mb.SO,ylab="Y");qqline(mb.SO,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SO, family="bcPower")$lambda# round(0.7159533) => y^(3/4)
tmb.SO <- mb.SO^(3/4)
plot(density(tmb.SO),main="Strikeouts");rug(tmb.SO)
plot(tmb.SO,StanRes1,xlab="Strikeouts",ylab="Standardized Residuals");abline(lsfit(tmb.SO,StanRes1),lty=2,col=2)
qqnorm(tmb.SO,ylab="Y");qqline(tmb.SO,lty=2,col=2)

#STOLEN BASES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SB <- lm.smk$TEAM_BASERUN_SB
m1 <- lm(mW~mb.SB, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SB),main="Stolen Bases");rug(mb.SB)
plot(mb.SB,StanRes1,xlab="Stolen Bases",ylab="Standardized Residuals");abline(lsfit(mb.SB,StanRes1),lty=2,col=2)
qqnorm(mb.SB,ylab="Y");qqline(mb.SB,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SB, family="bcPower")$lambda# round(-0.03916886) => y^(-1/25)
tmb.SB <- mb.SB^(-1/25)
plot(density(tmb.SB),main="Stolen Bases");rug(tmb.SB)
plot(tmb.SB,StanRes1,xlab="Stolen Bases",ylab="Standardized Residuals");abline(lsfit(tmb.SB,StanRes1),lty=2,col=2)
qqnorm(tmb.SB,ylab="Y");qqline(tmb.SB,lty=2,col=2)

#PITCHING HITS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.H <- lm.smk$TEAM_PITCHING_H
m1 <- lm(mW~mp.H, data = lm.smk)
plot(density(mp.H),main="Pitching Hits");rug(mp.H)
plot(mp.H,StanRes1,xlab="Pitching Hits",ylab="Standardized Residuals");abline(lsfit(mp.H,StanRes1),lty=2,col=2)
qqnorm(mp.H, ylab = "Y");qqline(mp.H, lty = 2, col=2)
##TRANSFORMATION
powerTransform(mp.H,family="bcPower")$lambda# round(-3.097364) => y^(-3)
tmp.H <- mp.H^(-3)
plot(density(tmp.H),main="Pitching Hits");rug(tmp.H)
plot(tmp.H,StanRes1,xlab="Pitching Hits",ylab="Standardized Residuals");abline(lsfit(tmp.H,StanRes1),lty=2,col=2)
qqnorm(tmp.H,ylab="Y");qqline(tmp.H,lty=2,col=2)

#PITCHING WALKS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.BB <- lm.smk$TEAM_PITCHING_BB
m1 <- lm(mW~mp.BB, data = lm.smk)
plot(density(mp.BB),main="Pitching Walks");rug(mp.BB)
plot(mp.BB,StanRes1,xlab="Pitching Walks",ylab="Standardized Residuals");abline(lsfit(mp.BB,StanRes1),lty=2,col=2)
qqnorm(mp.BB, ylab = "Y");qqline(mp.BB,lty = 2,col=2)
##TRANSFORMATION
powerTransform(mp.BB, family="bcPower")$lambda# round(0.1609713) => y^(1/6)
tmp.BB <- mp.BB^(1/6)
plot(density(tmp.BB),main="Pitching Walks");rug(tmp.BB)
plot(tmp.BB,StanRes1,xlab="Pitching Walks",ylab="Standardized Residuals");abline(lsfit(tmp.BB,StanRes1),lty=2,col=2)
qqnorm(tmp.BB,ylab="Y");qqline(tmp.BB,lty=2,col=2)

#PITCHING STRIKEOUTS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.SO <- lm.smk$TEAM_PITCHING_SO
m1 <- lm(mW~mp.SO, data = lm.smk)
plot(density(mp.SO),main="Pitching Strikeouts");rug(mp.SO)
plot(mp.SO,StanRes1,xlab="Pitching Strikeouts",ylab="Standardized Residuals");abline(lsfit(mp.SO,StanRes1),lty=2,col=2)
qqnorm(mp.SO, ylab = "Y");qqline(mp.SO,lty = 2,col=2)
##TRANSFORMATION
powerTransform(mp.SO, family="bcPower")$lambda#round(0.6522561) => y^(2/3)
tmp.SO <- mp.SO^(2/3)
plot(density(tmp.SO),main="Pitching Strikeouts");rug(tmp.SO)
plot(tmp.SO,StanRes1,xlab="Pitching Strikeouts",ylab="Standardized Residuals");abline(lsfit(tmp.SO,StanRes1),lty=2,col=2)
qqnorm(tmp.SO,ylab="Y");qqline(tmp.SO,lty=2,col=2)

#FIELDING ERRORS:
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.E <- lm.smk$TEAM_FIELDING_E 
m1 <- lm(mW~mf.E, data = lm.smk)
plot(density(mf.E),main="Fielding Errors");rug(mf.E)
plot(mf.E,StanRes1,xlab="Fielding Errors",ylab="Standardized Residuals");abline(lsfit(mf.E,StanRes1),lty=2,col=2)
qqnorm(mf.E, ylab = "Y");qqline(mf.E, lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.E,family="bcPower")$lambda#round(-0.939657) => (-9/10)
tmf.E <- mf.E^(-(9/10))
plot(density(tmf.E),main="Fielding Errors");rug(tmf.E)
plot(tmf.E,StanRes1,xlab="Fielding Errors",ylab="Standardized Residuals");abline(lsfit(tmf.E,StanRes1),lty=2,col=2)
qqnorm(tmf.E,ylab="Y");qqline(tmf.E,lty=2,col=2)

#DOUBLE PLAYS:
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.DP <- lm.smk$TEAM_FIELDING_DP
m1 <- lm(mW~mf.DP, data = lm.smk)
plot(density(mf.DP),main="Fielding Doubleplays");rug(mf.DP)
plot(mf.DP,StanRes1,xlab="Fielding Doubleplays",ylab="Standardized Residuals");abline(lsfit(mf.DP,StanRes1),lty=2,col=2)
qqnorm(mf.DP, ylab = "Y");qqline(mf.DP,lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.DP,family="bcPower")$lambda# round(1.49645) => 1.5
tmf.DP <- mf.DP^(3/2)
plot(density(tmf.DP),main="Fielding Doubleplays");rug(tmf.DP)
plot(tmf.DP,StanRes1,xlab="Fielding Doubleplays",ylab="Standardized Residuals");abline(lsfit(tmf.DP,StanRes1),lty=2,col=2)
qqnorm(tmf.DP,ylab="Y");qqline(tmf.DP,lty=2,col=2)

#FIELDING YIELD
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.FY <- lm.smk$TEAM_FIELDING_E + lm.smk$TEAM_FIELDING_DP * 2
m1 <- lm(mW~mf.FY, data = lm.smk)
plot(density(mf.FY),main="Fielding Yield");rug(mf.FY)
plot(mf.FY,StanRes1,xlab="Fielding Yield",ylab="Standardized Residuals");abline(lsfit(mf.FY,StanRes1),lty=2,col=2)
qqnorm(mf.FY, ylab = "Y");qqline(mf.FY,lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.FY,family="bcPower")$lambda# round(-2.066982) => -2
tmf.FY <- mf.FY^(-2)
plot(density(tmf.FY),main="Fielding Yield");rug(tmf.FY)
plot(tmf.FY,StanRes1,xlab="Fielding Yield",ylab="Standardized Residuals");abline(lsfit(tmf.FY,StanRes1),lty=2,col=2)
qqnorm(tmf.FY,ylab="Y");qqline(tmf.FY,lty=2,col=2)

par(mfrow=c(1,1))
```

##### Evaluate Correlations  
Evaluate Correlation between predictors so as to not introduce collinearity into the model.
```{r correlation, eval=TRUE, echo=FALSE}
# assign model one "lm.1" to data set with all NA's imputed and "bad" leverage points removed
par(cex = 0.65)
corrplot(
    cor(lm.smk[c(1:13)]), 
    type='lower', 
    tl.srt=45,
    addshade="positive",
    addCoef.col = rgb(0,0,0,alpha=0.3),
    addCoefasPercent = TRUE)
nrow(lm.smk)
```

##### Model Selection Strategy
Two common strategies for adding or removing variables in a multiple regression model are
called backward elimination and forward selection. These techniques are often referred to
as stepwise model selection strategies, because they add or delete one variable at a time
as they step through the candidate predictors.  Model 1 uses the forward selection strategy 
which adds variables two-at-a-time until variables cannot be found that improve the model 
as measured by adjusted $R^2$.
Diez, D.M., Barr, C.D., & etinkaya-Rundel, M. (2015). OpenIntro Statistics (3rd Ed). pg. 378

#####Start with p.Hits & p.Walks
```{r m1-iteration1,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Hits <- tmp.H 
p.Walks <- tmp.BB
m1 <- lm(Wins ~ p.Hits+p.Walks)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Hits+p.Walks)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

##DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Hits,StanRest,ylab="Standardized Residuals")
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Add b.Singles & b.Doubles
```{r m1-interation2,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Hits <- tmp.H 
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
m1 <- lm(Wins ~ p.Hits+p.Walks+b.Singles+b.Doubles)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Hits+p.Walks+b.Singles+b.Doubles)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values of p.Hits >  0.05 so it gets removed
```

#####Removed p.Hits
```{r m1-interation3,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5 and adjusted $R^2$ increased

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
##Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Added Stolen Bases and Double Plays
```{r m1-interation4,eval=TRUE,include=FALSE}
nrow(lm.smk)
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
#If plot doesn't shows a linear relationship with no pattern or skew the model lacks normality.

# normal distribution, and uniform distribution of residuals
```

#####Added b.Walks and p.Strikeouts
```{r m1-interation5,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
b.Walks <- tmb.BB
p.StrikeOuts <- tmp.SO
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.Walks+p.StrikeOuts)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.Walks+p.StrikeOuts)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 but VIFs  > 5
#highest vif is b.Walks so it gets removed
```

#####Remove b.Walks
```{r m1-interation6,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB 
f.DoublePlays <- tmf.DP
p.StrikeOuts <- tmp.SO
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+p.StrikeOuts)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+p.StrikeOuts)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-value for b.Walks > 0.05 all VIFs < 5
```

#####Add b.StrikeOuts, & b.Slugging
```{r m1-interation7,eval=TRUE,include=FALSE}
nrow(lm.smk)
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
b.StrikeOuts <- tmb.SO
b.Slugging <- tmb.SL
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.StrikeOuts+b.Slugging)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+b.StrikeOuts+b.Slugging)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)

#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(b.StrikeOuts,StanRest,ylab="Standardized Residuals")
plot(b.Slugging,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Add b.Fielding
\[
\begin{aligned}
\widehat{wins} &= \hat{\beta}_0 + 
                    \hat{\beta}_1 \times p.Walks 
                    \hat{\beta}_2 \times b.Singles + 
                    \hat{\beta}_3 \times b.Doubles +
                    \hat{\beta}_4 \times b.Stolen Bases +                    
                    \end{aligned}
\]
\[
\begin{aligned}
                    \hat{\beta}_5 \times f.Double Plays +  
                    \hat{\beta}_6 \times b.Strike Outs +
                    \hat{\beta}_7 \times b.Slugging +   
                    \hat{\beta}_8 \times b.Fielding Yield +                       
                    \end{aligned}
\]
nrow(lm.smk)
```{r m1-interation8,eval=TRUE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
Index <- lm.smk$INDEX 
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
b.StrikeOuts <- tmb.SO
b.Slugging <- tmb.SL
f.Fielding <- tmf.FY

m1 <- lm(Wins ~ -Index+p.Walks+b.Singles+b.Doubles+b.StolenBases+b.StrikeOuts+b.Slugging+f.Fielding)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.StolenBases+b.StrikeOuts+b.Slugging+f.Fielding)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)

#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(b.StrikeOuts,StanRest,ylab="Standardized Residuals")
plot(b.Slugging,StanRest,ylab="Standardized Residuals")
plot(f.Fielding,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

```{r}
#pred_eval.m1 <- round(predict(m1, eval_data))
#eval.BSO.imp <- impute(eval_data$TARGET_WINS, pred_eval.m1)
```


# Part 4. Select Models

```{r part4}
#########################################################################################
# 
# This file loads our preferred predictive model (TOTAL BASES PLUS)
# and uses that model to predict the TARGET_WINS variable of the MLB Evaluation 
# data set. 
#
# When finished, two separate files are written to a local hard disk directory:
#
# - one containing the entire EVALUATION data set after the TARGET_WINS variable has
#   been updated with the predicted values for each record;
#
# - one containing ONLY the INDEX and TARGET_WINS variables from the EVALUATION data set
#
# - NO screen output is generated at all by this code
#
#########################################################################################

# --------------------------------------------------------------
# read clean data set from Github

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  

# ---------------------------------------------------------------

# Build a model with Total Bases + SB + BB added and all of the other hitting vars removed

# create new variable and drop its components
mb_t <- mb_clean

mb_t$TB_PLUS <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR) + 
                       mb_clean$TEAM_BATTING_BB + mb_clean$TEAM_BASERUN_SB

# par(mfrow = c(1,1))
# hist(mb_t$TB_PLUS, breaks = 200)

# now drop 1B, 2B, 3B, HR, BB, SB
mb_tbp <- mb_t[,c(1, 2, 7, 9, 10, 11, 12, 13, 15)]


# -----------------------------------------------------------------------------------------

# REMOVE OUTLIERS AND REFIT

# Per Cooks Distance, remove items 836, 821, 1920, 1737, 1515


############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_tbp[-c(836, 821, 1920, 1737, 1515),]

# save first data set
mb_tbp_orig <- mb_tbp

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)

# keep the clean data set pure
mb_tbp <- mb_rem

# -------------------------------------------------------------------------------------------

## Now try same model but with FIELD_E transformed using Box-Cox

# TEAM_FIELDING_E: Box-cox says -1 power transform => 1/y 

mb_tbp$TEAM_FIELDING_E <- 1/mb_tbp$TEAM_FIELDING_E

# Now refit first model from above: Start with all variables
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_H - TEAM_PITCHING_SO)

# summary(model.4)
```

```{r}
# Now load evaluation data set and predict TARGET WINS

# load EVAL data set
eval.d <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-EvalData-.csv") 

# save original data
eval.2 <- eval.d

# creaet TB_PLUS and drop component variables
eval.2$TB_PLUS <- eval.2$TEAM_BATTING_1B + (2 * eval.2$TEAM_BATTING_2B) + 
                       (3 * eval.2$TEAM_BATTING_3B) + (4 * eval.2$TEAM_BATTING_HR) + 
                       eval.2$TEAM_BATTING_BB + eval.2$TEAM_BASERUN_SB

# par(mfrow = c(1,1))
# hist(eval.d$TB_PLUS, breaks = 30)

# now drop 1B, 2B, 3B, HR, BB, SB
eval.2 <- eval.2[,c(1, 6, 8, 9, 10, 11, 12, 14, 15)]

# transform TEAM_FIELDING_E using 1/y
eval.2$TEAM_FIELDING_E <- 1/eval.2$TEAM_FIELDING_E


# now predict TARGET_WINS using model.4
pred.TW <- round(predict(model.4, eval.2))

# add predicted variables to TARGET_WINS variable
eval.2$TARGET_WINS <- pred.TW
eval.d$TARGET_WINS <- pred.TW
```

```{r writecsv3, eval=FALSE}
# write entire updated EVAL data set to a CSV
write.csv(eval.d, file = "C:/SQLData/HW1-PRED-EVAL-ALLDATA.csv", row.names = FALSE)

# write full model EVAL data to a CSV file
write.csv(eval.d, file = "C:/SQLData/HW1-PRED-EVAL-ALL_M_DATA.csv", row.names = FALSE)

# now write just INDEX and TARGET_WINS to a separate file
eval.3 <- eval.2[,c(1,8)]

write.csv(eval.3, file = "C:/SQLData/HW1-PRED-EVAL-WINS-ONLY.csv", row.names = FALSE)

# end

```

```{r}
# clean up objects in memory
rm(list = ls())
```



# Conclusion

