---
title: 'Data 621 Homework 1: Moneyball'
author: Critical Thinking Group 2 - Armenoush Aslanian-Persico, James Topor, Jeff
  Nieman, Scott Karr
output:
  html_document:
    highlight: pygments
    keep_md: yes
    theme: cerulean
    toc: yes
  pdf_document: default
  word_document: default
---

```{r, eval=FALSE, include=FALSE, echo=FALSE}
BB.list <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/moneyball-training-data.csv", header=TRUE,
    stringsAsFactors=FALSE, sep=",")
attach(BB.list)

# describe structure of dataset
str(BB.list)

# plot(BB.list[,3:17])

mb_cor <- cor(BB.list[,3:17])

round(mb_cor, 3)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
```

__--------------------------------------------------------------------------------------------------------------------------__

# Part 1: Data Exploration

## Data Summary 
The original training set data provided included 17 attributes with 2277 observations.  

TRAINING DATASET:  Descriptive Statistics Prior to "cleaning" data

|variables       | n     | mean | sd   | med  | min  | max   | range |  skew | kurtosis | se    | NAs
|--------------  | ----- | ---- | ---- | ---- | ---- | ----- | ----- | ----- | -------  | ----- | ---
|INDEX           | 2276  | 1268 |  736 | 1271 |    1 | 2535  |  2534 |  0.00 |  -1.22   | 15.43 |     
|TARGET_WINS     | 2276  |   81 |   16 |   82 |    0 |   146 |   146 | -0.40 |   1.03   |  0.33 |    
|TEAM_BATTING_H  | 2276  | 1470 |  145 | 1454 |  891 |  2554 |  1663 |  1.57 |   7.28   |  3.03 |    
|TEAM_BATTING_2B | 2276  |  241 |   47 |  238 |   69 |   458 |   389 |  0.22 |   0.01   |  0.98 |     
|TEAM_BATTING_3B | 2276  |   55 |   28 |   47 |    0 |   223 |   223 |  1.11 |   1.50   |  0.59 |    
|TEAM_BATTING_HR | 2276  |  100 |   61 |  102 |    0 |   264 |   264 |  0.19 |  -0.96   |  1.27 |    
|TEAM_BATTING_BB | 2276  |  502 |  123 |  512 |    0 |   878 |   878 | -1.03 |   2.18   |  2.57 |     
|TEAM_BATTING_SO | 2174  |  736 |  240 |  750 |    0 |   697 |   697 |  1.97 |   5.49   |  1.90 |  131
|TEAM_BASERUN_CS | 1504  |   53 |   23 |   49 |    0 |   201 |   201 |  1.98 |   7.62   |  0.59 |  772
|TEAM_BATTING_HBP|  191  |   59 |   13 |   58 |   29 |    95 |    66 |  0.32 |  -0.11   |  0.94 | 2085
|TEAM_PITCHING_H | 2276  | 1779 | 1407 | 1518 | 1137 | 30132 | 28995 | 10.33 | 141.84   | 29.49 |     
|TEAM_PITCHING_HR| 2276  |  106 |   61 |  107 |    0 |   343 |   343 |  0.29 |  -0.60   |  1.28 |      
|TEAM_PITCHING_BB| 2276  |  553 |  166 |  536 |    0 |  3645 |  3645 |  6.74 |  96.97   |  3.49 |      
|TEAM_PITCHING_SO| 2174  |  818 |  553 |  813 |    0 | 19278 | 19278 | 22.17 | 671.19   | 11.86 |  102
|TEAM_FIELDING_E | 2276  |  246 |  228 |  159 |   65 |  1898 |  1833 |  2.99 |  10.97   |  4.77 |      
|TEAM_FIELDING_DP| 1990  |  146 |   26 |  149 |   52 |   228 |   176 | -0.39 |   0.18   |  0.59 |  286




## Data Plots
Bar Chart or Box Plot of the data

## Correlation Plot
Is the data correlated to the target variable (or to other variables?)

__--------------------------------------------------------------------------------------------------------------------------__

#Part 2: Data Preparation

Our data preparation efforts for the training data set consisted of the creation of one new derived variable, removing four predictor variables, imputing values for the remaining variables that had missing values (NA's), and removal of a relatively small number of records that contained clearly egregious outlier values for particular variables. The results of these efforts were subsequently used as the basis for each of the five different linear models we created and evaluated. 

## New Variable Creation

We began our data preparation efforts by creating a new variable TEAM_BATTING_1B which represents offensive single base hits. The variable was created by subtracting out the TEAM_BATTING doubles, triples and home runs from the TEAM_BATTING_H variable. 

The TEAM_BATTING_H variable was then dropped from the data set. We believe that separating out singles from the other unique hit values will minimize collinearity.

## Removal of Variables Due to Collinearity and/or Missing Values

The results of our data exploration efforts led us to drop three other variables from the data set:

1. __TEAM_BATTING_HBP__: The TEAM_BATTING_HBP variable has very little correlation with the TARGET_WINS response variable and also has 2085 missing values out of a total of 2277. Since it would be very difficult to accurately impute such a large proportion of any variable's missing values, we chose to exclude the variable from our analysis.

2. __TEAM_BASERUN_CS__: This variable is strongly correlated (65.5%) with the TEAM_BASERUN_SB variable and is the 2nd largest source of NA's in our data set. These combined facts led us to exclude the variable from our analysis.

3. __TEAM_PITCHING_HR__: This variable is 97% correlated with TEAM_BATTING_HR. In fact, 815 cases (more than 35% of our total cases) have *IDENTICAL* values for pitched and batted HR's.  This high degree of correlation may be due to the time series nature of the data: as baseball evolved, more home runs were hit, which naturally causes the number of pitched home runs to increase. The statistics are basically opposite sides of the same coin to a large degree (even if there may be some variability between individual teams in any given year). The fact that these two variables are nearly perfectly correlated indicates that one of them can legitimately be removed from the data set, and we chose __TEAM_PITCHING_HR__ since we believe the batting HR metric will be more predictive of TARGET_WINS than will the pitching HR statistic.

## Imputation of Missing Values (NA's)

### Filling Missing values in the Training Data Set

We then worked on imputing the remaining missing data. To do this we made use of a linear regression approach recommended by Faraway (p.201) and Fox (p.611). We decided against the using the mean or median as a replacement value for NA's since regression yields imputed values that are much more consistent with the actual distribution of the data while introducing much less potential bias than use of a mean or median would incur.

We built regression imputation models for the variables listed below. Each model was analyzed to ensure that there were no collinearity issues and all p-values were < $.05$. While we did not not have the time or resources to run a complete cadre of residual diagnostics on these imputation models, we did verify that their results produced imputed distributions of the subject variables that were consistent with those of the original NA-populated data. It is our belief that this consistency indicates that the resulting predicted values for the missing values are better to use than simply filling the NA's with a mean or median.

The variables we built imputation regression models for are described below. The R code for each of the regression models is contained in the Appendix.

1. __TEAM_BATTING_SO__: The adjusted $R^2$ value for this regression model was 0.7223 and, after populating the NA's using the imputed values, yielded a distribution that matched that of the variable prior to the NA's being filled. For this variable we imputed a total of 102 missing values via regression.

2. __TEAM_PITCHING_SO__: The adjusted $R^2$ value for this regression model was 0.9952. A total of 102 missing values were imputed via regression for this variable.

3. __TEAM_BASERUN_SB__: The adjusted $R^2$ value for this regression model was 0.3427. Despite the adjusted $R^2$ being low relative to the models described above, after the model yielded a distribution that matched that of the variable prior to the NA's being filled. A total of 131 missing values were imputed via regression for this variable.

4. __TEAM_FIELDING_DP__: The adjusted $R^2$ for this model was 0.3904. A total of 286 missing values were imputed and the resulting distribution matched that of the variable prior to the NA's being filled. 

### Filling Missing Values in the Evaluation Data Set

These four imputation regression models were then used to populate missing values for the same variables within the evaluation data set. Clearly, any models we might have constructed for purposes of predicting WINS for the evaluation set would have failed to work on any records within that data set that contained NA's. Applying the imputation models to the evaluation data allows us to avoid that problem.

## Removal of Extreme Outliers From Data Sets

### Extreme Outlier Removal: Training Data

We completed our data preparation efforts by eliminating some clearly egregious outliers identified via research through *baseball-almanac.com*. This approach is suggested by Sheather (p. 57).  For example, the record for the most pitching strikeouts in a single season is 1450 by the 2014 Cleveland Indians. Therefore we know that any records having TEAM_PITCHING_SO values above that point is an aberration, and any records containing such values were summarily removed from the data set. 

Similarly, the most errors by team in a single season was 639 by Philadelphia in 1883. Prorating to 162 games we calculated that we should discard any records containing TEAM_FIELDING_E values above 1046. 

The TEAM_PITCHING_H variable also appeared to have numerous egregious outliers. For example, the most offensive hits by a team in a single season was 1730. As such, it is highly unlikely that any pitching staff would surrender more than 3000 hits in a single season since such a total would indicate the team was allowing more than 18 hits per game. As such, any records having a TEAM_PITCHING_H value > 3000 was removed from the data set.

Removing the egregious outliers resulting in a sum total of 104 records being removed from the training data set. Removal of these outliers helped to normalize our data and thereby improve the expected performance of our linear models.

### Extreme Outlier Removal: Evaluation Data

Similar extreme outliers were subsequently removed from the evaluation data set since it is unlikely that any regression models we might have built would have generated realistic predictive results for records containing such outliers. Of the 259 records contained in the evaluation data set, a total of 11 were removed by this process. 

__*As such, our Evaluation data set is comprised of a total of 248 records rather than the 259 records provided in the original data*__.

Our Evaluation data set with the NA's filled and outliers removed can be found here:

https://github.com/spsstudent15/2016-02-621-W1/blob/master/621-HW1-Clean-EvalData-.csv

## Other Data Prepation Transformations: Refer to Model Descriptions

Other *model-specific* data transformations were also used, including use of Box-Cox power transforms and linear combinations of variables. These model-specific transformations are discussed within the individual model writeups provided in __Part 3__.

__--------------------------------------------------------------------------------------------------------------------------__

# Part 3: Build Models

## Model 1 - General Model Using Backward Selection

This model applies simple Backward Selection methods through the use of p-values and variance inflation factors (VIF) against the following predictor variables:

- TEAM_BATTING_1B (derived variable)  
- TEAM_BATTING_2B  
- TEAM_BATTING_3B  
- TEAM_BATTING_HR  
- TEAM_BATTING_BB  
- TEAM_BATTING_SO  
- TEAM_BASERUN_SB  
- TEAM_PITCHING_H  
- TEAM_PITCHING_BB  
- TEAM_PITCHING_SO  
- TEAM_FIELDING_E  
- TEAM_FIELDING_DP  

Simply removing the *TEAM_BATTING_1B* variable yielded a model with all p-values less than $.05$. However, VIF analysis showed evidence of multiple collinear variables within the model. Subsequent removals of *TEAM_PITCHING_SO* and *TEAM_PITCHING_BB* due to collinearity yielded a model that called for the removal of *TEAM_BATTING_2B* on the basis of its p-value.

The final model of that iteration of the linear modeling process showed clear evidence of a number of outliers as evidenced in R's summary diagnostic plots. Those outliers were removed via a series of additional iterations yielding the following final model, which varies from the inital iteration in that it includes *TEAM_BATTING_2B* due to the fact that removing the outliers improved the statistical significance of the variable :


| Coefficient   | Variable  
| ------------  | ------------
| 66.261        | Intercept 
| - 0.017       | TEAM_BATTING_2B
| + 0.150       | TEAM_BATTING_3B
| + 0.109       | TEAM_BATTING_HR
| + 0.022       | TEAM_BATTING_BB
| - 0.019       | TEAM_BATTING_SO
| + 0.065       | TEAM_BASERUN_SB
| + 0.016       | TEAM_PITCHING_H
| - 0.075       | TEAM_FIELDING_E
| - 0.109       | TEAM_FIELDING_DP  



| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.49 | 0.3598  | 0.3572    | 134.4   | 132 

                     
However, the diagnostic plots of that model showed a lack of linearity between the response variable TARGET_WINS and the predictor variable TEAM_FIELDING_E as evidenced in the Added Variable plots shown in the Appendix. Furthermore, the plots of standardized residuals against each of the predictor variables showed evidence of non-constant variability for variables such as TEAM_BATTING_HR, TEAM_BATTING_SO, TEAM_BASERUN_SB, and TEAM_FIELDING_E.

The TEAM_FIELDING_E variable was subsequently transformed using a Box-Cox recommended power transform of (-1), or (1/y) and the model was re-run. The resulting Added Variable plots showed that all predictors are linearly related to the response, and we see an improvement in the variability of the residuals relative to TEAM_FIELDING_E. Furthermore, the plot of Y against the fitted values showed an improvement in the linearity of the model. 

Therefore, this model appears to be an improvement over the first model when the residual plots are considered. The characteristic equation indicated by the model is as follows:


| Coefficient   | Variable  
| ------------  | ------------
| 52.88         | Intercept 
| + 0.168       | TEAM_BATTING_3B
| + 0.096       | TEAM_BATTING_HR
| + 0.027       | TEAM_BATTING_BB
| - 0.027       | TEAM_BATTING_SO
| + 0.034       | TEAM_BASERUN_SB
| + 0.004       | TEAM_PITCHING_H
| + 3252.31     | 1/TEAM_FIELDING_E
| - 0.102       | TEAM_FIELDING_DP


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.86 | 0.3168  | 0.3143    | 124.8   | 141


The coefficients for TEAM_BATTING_3B, TEAM_BATTING_HR, TEAM_BATTING_BB, TEAM_BATTING_SO, and TEAM_BASERUN_SB all make sense intuitively. The TEAM_FIELDING_DP coefficient is less intuitive since one would expect more defensive double plays to improve a team's chances of winning games. However, the variable itself is *negatively* correlated with TARGET_WINS as shown in the __Data Exploration__ section above. As such, we shouldn't be surprised to see a negative coefficient for it here. Similarly, the coefficient for TEAM_PITCHING_H is also counterintuitive, but the variable is actually positively correlated with TARGET_WINS as shown in the __Data Exploration__ section. Finally, TEAM_FIELDING_E has changed from negative in the earlier model to positive here. However, this is due to the fact that the coefficient now applies to the *transformed* version of the variable rather than the nominal values of the variable.

While this model is an improvement over the initial model, we still have component variables that appear to lack constant variability relative to the residuals for variables such as TEAM_BASERUN_SB. The lack of constant variability in the residuals is likely related to the skewed nature of the distributions of those individual variables. 

In the other models discussed herein we attempt to address some of the skew issues via various methods, including Box-Cox recommended power transforms and linear combinations of various variables.


__--------------------------------------------------------------------------------------------------------------------------__

## Model 2 - Total Bases 

This model attempts to address some of the lack of constant variability found in the "General Model" discussed above by employing a linear combination of four of the predictor variables to calculate the baseball statistc known as "Total Bases". Total Bases is calculated using what our data set refers to as "TEAM_BATTING" variables as follows:

- Singles + (2 * Doubles) + (3 * Triples) = (4 * Home Runs)

Inclusion of this new variable allows us to eliminate the four component variables from the model. In fact, the TOTAL_BASES variable appears to be normally distributed, thereby negating the skew issues that were evident with its component variables.

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# read clean data set from Github

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  

# create new variable and drop its components
mb_t <- mb_clean

mb_t$TB_ <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR)

# set smaller margins to reduce size of histogram
par(mai=c(.5,.5,.5,.5))
par(mfrow = c(1, 1))
hist(mb_t$TB, breaks = 200, main = "Distribution of TOTAL_BASES Variable", xlab = "TOTAL_BASES")

# set margins back to default
par(mai=c(1.02,0.82,0.82,0.42))
rm(mb_clean)
rm(mb_t)

```

This model applies simple Backward Selection methods through the use of p-values and variance inflation factors (VIF) against the following predictor variables:

- TEAM_TOTAL_BASES (derived variable)  
- TEAM_BATTING_BB  
- TEAM_BATTING_SO  
- TEAM_BASERUN_SB  
- TEAM_PITCHING_H  
- TEAM_PITCHING_BB  
- TEAM_PITCHING_SO  
- TEAM_FIELDING_E  
- TEAM_FIELDING_DP  

Three iterations of p-value / VIF backward selection removed TEAM_PITCHING_SO and TEAM_PITCHING_BB from the model. All other variables remained statistically significant with no signficant collinearity. However, evidence of multiple outliers was found via R's summary diagnostic plots. Those outliers were removed via a series of additional iterations yielding the following final model:


| Coefficient   | Variable  
| ------------  | ------------
| 48.486        | Intercept 
| + 0.022       | TEAM_BATTING_BB
| - 0.015       | TEAM_BATTING_SO
| + 0.063       | TEAM_BASERUN_SB
| + 0.010       | TEAM_PITCHING_H
| - 0.064       | TEAM_FIELDING_E
| - 0.117       | TEAM_FIELDING_DP
| + 0.018       | TOTAL_BASES


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.7  | 0.3365  | 0.3343    | 156     | 137


The diagnostic plots of that model show a lack of linearity between the response variable TARGET_WINS and the predictor variable TEAM_FIELDING_E as evidenced in the Added Variable plots shown in the Appendix. Furthermore, the plots of standardized residuals against each of the predictor variables showed evidence of non-constant variability for variables such as TEAM_BATTING_SO, TEAM_BASERUN_SB, and TEAM_FIELDING_E. 

The TEAM_FIELDING_E variable was subsequently transformed using a Box-Cox recommended power transform of (-1), or (1/y) and the model was re-run. The resulting Added Variable plots show that all predictors are linearly related to the response, and we see an improvement in the variability of the residuals relative to TEAM_FIELDING_E. Furthermore, the plot of Y against the fitted values shows an improvement in the linearity of the model. 

Therefore, this model appears to be an improvement over the first TOTAL_BASES model when the residual plots are considered. The characteristic equation indicated by the model is as follows:

| Coefficient   | Variable  
| ------------  | ------------
| 39.164        | Intercept 
| + 0.025       | TEAM_BATTING_BB
| - 0.025       | TEAM_BATTING_SO
| + 0.038       | TEAM_BASERUN_SB  
| + 2714.54     | 1/TEAM_FIELDING_E
| - 0.115       | TEAM_FIELDING_DP
| + 0.0197       | TOTAL_BASES


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.97 | 0.3048  | 0.3029    | 157.5   | 143




The coefficients for TEAM_BATTING_BB, TEAM_BATTING_SO, TEAM_BASERUN_SB, and TOTAL_BASES all make sense intuitively. The TEAM_FIELDING_DP coefficient is less intuitive since one would expect more defensive double plays to improve a team's chances of winning games. However, the variable itself is *negatively* correlated with TARGET_WINS as shown in the __Data Exploration__ section above. As such, we shouldn't be surprised to see a negative coefficient for it here. TEAM_FIELDING_E has changed from negative in the earlier model to positive here. However, this is due to the fact that the coefficient now applies to the *transformed* version of the variable rather than the nominal values of the variable.

__--------------------------------------------------------------------------------------------------------------------------__

## Model 3 - Total Bases PLUS

This model attempts to improve upon the results of the "Total Bases" model by extending the TOTAL_BASES variable to include the TEAM_BATTING_BB and TEAM_BASERUN_SB variables. The logic behind adding these two variables to the TOTAL_BASES variable comes from the fact that both, like the component variables of TOTAL_BASES, represent basepath advancements by a team's offense.

"Total Bases Plus"" (referred to as TB_PLUS hereon) is calculated using what our data set refers to as "TEAM_BATTING" and "TEAM_BASERUN" variables as follows:

- Singles + (2 * Doubles) + (3 * Triples) = (4 * Home Runs) + BB + SB  

Inclusion of this new variable allows us to eliminate the two additional component variables from the model. In fact, the TB_PLUS variable, like the TOTAL_BASES variable used earlier appears to be normally distributed, thereby negating the skew issues that were evident with its component variables. A histogram of the distribution of the derived TB_PLUS variable is shown below:


```{r, echo = FALSE, warning=FALSE, message=FALSE}
# read clean data set from Github

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  

# ---------------------------------------------------------------

# Build a model with Total Bases + SB + BB added and all of the other hitting vars removed

# create new variable and drop its components
mb_t <- mb_clean

mb_t$TB_PLUS <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR) + 
                       mb_clean$TEAM_BATTING_BB + mb_clean$TEAM_BASERUN_SB

par(mfrow = c(1,1))
hist(mb_t$TB_PLUS, breaks = 200, main = "Distribution of TB_PLUS Variable", xlab = "TB_PLUS")

rm(mb_clean)
rm(mb_t)

```

This model applies simple Backward Selection methods through the use of p-values and variance inflation factors (VIF) against the following predictor variables:

- TB_PLUS (derived variable)  
- TEAM_BATTING_SO  
- TEAM_PITCHING_H  
- TEAM_PITCHING_BB  
- TEAM_PITCHING_SO  
- TEAM_FIELDING_E  
- TEAM_FIELDING_DP  

Four iterations of p-value / VIF backward selection removed TEAM_PITCHING_H, TEAM_PITCHING_SO and TEAM_PITCHING_BB from the model. All other variables remained statistically significant with no signficant collinearity. However, evidence of multiple outliers was found via R's summary diagnostic plots. Those outliers were removed via a series of additional iterations yielding the following final model:


| Coefficient   | Variable  
| ------------  | ------------
| 52.330        | Intercept 
| - 0.016       | TEAM_BATTING_SO
| - 0.034       | TEAM_FIELDING_E
| - 0.154       | TEAM_FIELDING_DP
| + 0.025       | TB_PLUS


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 12.12 | 0.2944  | 0.2931    | 225.5   | 145


The coefficients for TEAM_BATTING_SO and TEAM_FIELDING_E make sense intuitively: the more strikeouts a team's offense has, the less likely it is to put the ball in play, and the more fielding errors a team commits, the more likely they are to lose games. The TEAM_FIELDING_DP coefficient is less intuitive since one would expect more defensive double plays to improve a team's chances of winning games. However, the variable itself is *negatively* correlated with TARGET_WINS as shown in the __Data Exploration__ section above. As such, we shouldn't be surprised to see a negative coefficient for it here. Finally, the coefficient for TB_PLUS is postively correlated with the response variable, which shouldn't surprise us since it encapsulates all of a team's offense hits, stolen bases, and bases on balls.

As with the "General Model" and the "Total Bases" model, the diagnostic plots for this model showed a lack of linearity between the response variable TARGET_WINS and the predictor variable TEAM_FIELDING_E as evidenced in the Added Variable plots shown in the Appendix. Furthermore, the plots of standardized residuals against each of the predictor variables showed evidence of non-constant variability for the variables TEAM_BATTING_SO and TEAM_FIELDING_E.

The TEAM_FIELDING_E variable was subsequently transformed using a Box-Cox recommended power transform of (-1), or (1/y) and the model was re-run. The resulting Added Variable plots showed that all predictors are linearly related to the response, and we found an improvement in the variability of the residuals relative to TEAM_FIELDING_E. Furthermore, the plot of Y against the fitted values shows a non-skewed linear relationship. 

Therefore, this model appears to be an improvement over the first TB_PLUS model when the residual plots are considered. Furthermore, the number of predictor variables used here is two fewer than that of the "Total Bases" model discussed earlier. The characteristic equation indicated by the model is as follows:

                     
| Coefficient   | Variable  
| ------------  | ------------
| 42.160        | Intercept 
| - 0.023       | TEAM_BATTING_SO
| + 2366.82     | 1/TEAM_FIELDING_E
| - 0.140       | TEAM_FIELDING_DP
| + 0.022       | TB_PLUS


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 12.13 | 0.2932  | 0.2919    | 223.3   | 147


As we can see, the coefficient for TEAM_FIELDING_E has changed from negative to positive. However, this is due to the fact that the coefficient now applies to the *transformed* version of the variable rather than the nominal values of the variable. The other coefficients remain the same.

__--------------------------------------------------------------------------------------------------------------------------__

## Model 4 :Sabermetrics Model

Sabermetrics has become the rage in baseball, actually popularized by Billy Beane and the data set we are exploring.  As a result of this, we built a model that centers around one of these advance analyticsm known as BsR or base runs. This statistic was designed by David Smyth in the 1990's and estimates the amount of runs a team SHOULD score, which made a unique approach as the data set provided did not include runs (see http://tangotiger.net/wiki_archive/Base_Runs.html for more information). The formula is as follows:  

BSR = A*B/(B+C) +D where:  

- A = TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_BB  

- B = 1.02*(1.4*TEAM_TOTAL_BASES -0.6*TEAM_BATTING_H + 0.1*TEAM_BATTING_BB)  

- C = AT BATS - TEAM_BATTING_H (which we approximated with 3*TEAM_BATTING_H as the average batting average is around 0.250)  

- D = TEAM_BATTING_HR  
    
Since we eliminated the value of TEAM_BATTING_H we simply summed up singles, doubles, triples and home runs in the actual code, and the approach for TEAM_TOTAL_BASES is described in model 2.  The data for BSR exhibit a fairly normal distribution.

- __Insert BSR histogram here??__

Since BSR is a combination of all of the batting variables, we simply eliminated them and created a very strong model on the first iteration.  All p-values were very low, and the variation are all below 5 showing no probems with collinearity.  The characteristic equation indicated by the model is as follows:


| Coefficient   | Variable  
| ------------  | ------------
| 40.689953     | Intercept 
| + 0.062185    | BSR
| - 0.116618    | TEAM_FIELDING_DP
| - 0.058867    | TEAM_FIELDING_E
| + 0.60318     | TEAM_BASERUN_SB
| - 0.011451    | TEAM_PITCHING_SO
| + 0.019421    | TEAM_PITCHING_H
| - 0.017615    | TEAM_PITCHING_BB


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.99 | 0.3229  | 0.3207    | 147.4   | 144



The coefficients for this model overall make sense with one exception.  Errors and pitching walks contribute to fewer wins, and stolen bases and the BSR metric have strong influence on increasing wins.  Double plays do habve a slightly negative value, although this could be explained by a team allowing a large number of baserunners.     The positive impact of allowing pitching hits was puzzling.

__--------------------------------------------------------------------------------------------------------------------------__

## Model 5


- INSERT WRITEUP HERE

Placeholder tables - need to be filled in with ACTUAL metrics for Model 5

| Coefficient   | Variable  
| ------------  | ------------
| 52.88         | Intercept 
| + 0.168       | TEAM_BATTING_3B
| + 0.096       | TEAM_BATTING_HR
| + 0.027       | TEAM_BATTING_BB
| - 0.027       | TEAM_BATTING_SO
| + 0.034       | TEAM_BASERUN_SB
| + 0.004       | TEAM_PITCHING_H
| + 3252.31     | 1/TEAM_FIELDING_E
| - 0.102       | TEAM_FIELDING_DP


| RSE   | R^2     | Adj. R^2  | F Stat. | MSE
| ----- | ------- | --------- | ------- | -----
| 11.86 | 0.3168  | 0.3143    | 124.8   | 141


__--------------------------------------------------------------------------------------------------------------------------__

# Part 4. Select Models

The chart below summarizes the model statistics for all five of our models. The models are listed from left to right in accordance with the order in which they were described in __Part 3__ herein (Models 1-5). Please refer to the Appendix for detailed descriptions, diagnostic plots, and procedural details on the derivation of each of these models.


| Metric    | General Model | Total Bases | TB PLUS | Sabermetrics  | Model 5  
| --------- | ------------- | ----------- | ------- | ------------- | -------  
| RSE       | 11.86         | 11.97       | 12.12   | 11.99         | 0 
| R^2       | 0.3168        | 0.3048      | 0.2994  | 0.3229        | 0
| Adj. R^2  | 0.3143        | 0.3029      | 0.2931  | 0.3207        | 0
| F Stat.   | 124.8         | 157.5       | 224.3   | 147.4         | 0 
| MSE       | 141           | 143         | 147     | 144           | 0 


Each of our five models converged on similar $R^2$ values, RSE's, and MSE's, and all yielded residuals that were distributed normally without signficant evidence of highly leveraged outliers.  No signficant collinearity exists within any of the five models for any of their component predictor variables. 

Of the five, the General Model had the least favorable residual characteristics, with multiple predictors showing non-constant variability relative to the residuals, while having a relatively low F-statistic when compared to the other models.

Of the remaining four models, the Total Bases model was a great improvement over the General Model. However, it too displayed some lack of constant variability of residuals relative to a couple of the predictor variables.

The .... __DESCRIBE SCOTT's MODEL HERE__...

Of the remaining two (TB PLUS and Sabermetrics), while the Sabermetrics model yields a slightly large $R^2$ value, the TB Plus model is simpler in that it makes use of only 4 predictor variables while possessing a much larger F-statistic. Such a large difference in F-statistics indicates that the TB Plus model is explaining more of the variability of the training data than is the Sabermetrics model. 
  
Therefore, __we selected the TB PLUS model__ as the basis for our prediction of TARGET_WINS for the Evaluation data set. To ensure the model's efficacy when applied to the Evaluation data set, we first applied a set of transformations to that data set that were identical to those applied to the Training data set during the development of the TB PLUS model. The TB PLUS model was then applied to yield a set of INDEX / TARGET_WINS pairs.

Since displaying the full set of predicted values would consume a large number of pages herein, a sample of the first 10 rows of that data set is displayed below:

```{r, echo = FALSE, warning=FALSE, message=FALSE}

# read clean data set from Github

eval.w <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/HW1-PRED-EVAL-WINS-ONLY.csv")  

# display 10 rows
kable(head(eval.w, n = 10), alingn = 'l', padding = 0)

rm(eval.w)
```

The full set of predicted TARGET_WINS can be found at the following web link:

https://github.com/spsstudent15/2016-02-621-W1/blob/master/HW1-PRED-EVAL-WINS-ONLY.csv
