---
title: "Data 621 HW1"
author: "Jeff Nieman"
date: "June 9, 2016"
output: html_document
---
#####Creating a new column for batting singles and eliminating hits for batting
```{r}
mb_e <- read.csv("C:/Users/jenieman/Documents/CUNY/Data 621/moneyball-training-data.csv")  
#eliminate index column
mb_e1 <- mb_e[,-1]

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-2]
mb_e1 <- as.data.frame(mb_e1)
```

#####Building a regression model and filling in NA's for SB.  
Note:  This approach is suggested in LMAR p. 201.  "A more sophisticated alternative to mean imputation is to use regression methods to predict the missing values of the covariates."
```{r}
SB <- lm(data=mb_e1, TEAM_BASERUN_SB~.)
summary(SB)

#eliminate CS as there are no blank SB's with a value for CS + eliminate pitching, wins and fielding variables
SB1 <- lm(data=mb_e1, TEAM_BASERUN_SB~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB)
summary(SB1)

#eliminate singles
SB2 <- lm(data=mb_e1, TEAM_BASERUN_SB~TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB)
summary(SB2)

#fill in NA for SB
mb_e2 <- mb_e1
mb_e2$TEAM_BASERUN_SB[is.na(mb_e2$TEAM_BASERUN_SB)]<-round(79.36805-0.19419*mb_e2$TEAM_BATTING_2B+1.41686*mb_e2$TEAM_BATTING_3B-0.24513*mb_e2$TEAM_BATTING_HR+0.08060*mb_e2$TEAM_BATTING_BB)
```

#####Building a linear regression model and filling in NA's for CS
```{r}
CS <- lm(data=mb_e2, TEAM_BASERUN_CS~.)
summary(CS)

#eliminate wins, pitching and fielding 
CS1 <- lm(data=mb_e2, TEAM_BASERUN_CS~TEAM_BASERUN_SB +TEAM_BATTING_1B +TEAM_BATTING_2B + TEAM_BATTING_3B +TEAM_BATTING_BB + TEAM_BATTING_HR)
summary(CS1)

#eliminate walks
CS2 <- lm(data=mb_e2, TEAM_BASERUN_CS~TEAM_BASERUN_SB +TEAM_BATTING_2B + TEAM_BATTING_3B +TEAM_BATTING_1B + TEAM_BATTING_HR)
summary(CS2)

#fill in NA for CS
mb_e3 <- mb_e2
mb_e3$TEAM_BASERUN_CS[is.na(mb_e3$TEAM_BASERUN_CS)]<-round(49.356793+0.322543*mb_e3$TEAM_BASERUN_SB-0.044486*mb_e3$TEAM_BATTING_2B+0.281124*mb_e3$TEAM_BATTING_3B-0.10797*mb_e3$TEAM_BATTING_HR-0.014034*mb_e3$TEAM_BATTING_1B)

```

#####Building a regression model and filling in NA's for batting SO's
```{r}
BSO <- lm(data=mb_e3, TEAM_BATTING_SO~.)
summary(BSO)

#eliminate fielding and wins and baserunning and HBP and pitching SO's as it contains similar blanks
BSO1 <- lm(data=mb_e3, TEAM_BATTING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(BSO1)

#eliminate pitching HR's
BSO2 <- lm(data=mb_e3, TEAM_BATTING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_BB + TEAM_PITCHING_H  + TEAM_PITCHING_BB + TEAM_BATTING_HR)
summary(BSO2)

#fill in NA for batting SO
mb_e4 <- mb_e3
mb_e4$TEAM_BATTING_SO[is.na(mb_e4$TEAM_BATTING_SO)]<-round(1605-0.8434*mb_e4$TEAM_BATTING_1B+0.2832*mb_e4$TEAM_BATTING_2B-1.348*mb_e4$TEAM_BATTING_3B-0.3493*mb_e4$TEAM_BATTING_BB-0.02903*mb_e4$TEAM_PITCHING_H+0.1657*mb_e4$TEAM_PITCHING_BB+1.703*mb_e$TEAM_BATTING_HR)

```

#####Building a regression model and filling in NA's for pitching SO's
```{r}
PSO <- lm(data=mb_e4, TEAM_PITCHING_SO~.)
summary(PSO)

#eliminate wins, fielding, baserunning and HBP
PSO1 <- lm(data=mb_e4, TEAM_PITCHING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_BATTING_SO + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(PSO1)

#eliminate batting 3B's
PSO2 <- lm(data=mb_e4, TEAM_PITCHING_SO~TEAM_BATTING_1B + TEAM_BATTING_2B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_PITCHING_H + TEAM_BATTING_SO + TEAM_PITCHING_BB + TEAM_PITCHING_HR)
summary(PSO2)

#replace NA with values for pitching SO
mb_e5 <- mb_e4
mb_e5$TEAM_PITCHING_SO[is.na(mb_e5$TEAM_PITCHING_SO)]<-round(4422.87422-0.46455*mb_e5$TEAM_BATTING_1B + 0.72823*mb_e5$TEAM_BATTING_2B +8.06941*mb_e5$TEAM_BATTING_HR -3.45005*mb_e5$TEAM_BATTING_BB-0.01815*mb_e5$TEAM_PITCHING_H+1.20501*mb_e5$TEAM_BATTING_SO+3.03598*mb_e5$TEAM_PITCHING_BB-8.40807*mb_e$TEAM_PITCHING_HR)

```

#####Building a linear regression and filling in NA's for DP
```{r}


#build a regression model for DP's

DP <- lm(data=mb_e5, TEAM_FIELDING_DP~.)
summary(DP)

#eliminate wins, hitting, HBP

DP1 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_BASERUN_CS+TEAM_PITCHING_H+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP1)

#eliminate pitching hits allowed
DP2 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_BASERUN_CS+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP2)

#eliminate CS as it makes no sense
DP3 <- lm(data=mb_e5, TEAM_FIELDING_DP~ TEAM_BASERUN_SB+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_FIELDING_E)
summary(DP3)

#replace NA with values for DP
mb_e6 <- mb_e5
mb_e6$TEAM_FIELDING_DP[is.na(mb_e6$TEAM_FIELDING_DP)]<- round(158.8-0.1235*mb_e6$TEAM_BASERUN_SB +0.03320*mb_e6$TEAM_PITCHING_HR+0.02815*mb_e6$TEAM_PITCHING_BB - 0.006109*mb_e6$TEAM_PITCHING_SO - 0.06573*mb_e6$TEAM_FIELDING_E)

summary(mb_e6)
#only NA's left are HBP


