---
title: "621 HW 1 v 4"
author: "Jeff Nieman, Scott Karr, James Topor, Armenoush"
date: "June 13, 2016"
output: pdf_document
---

```{r load-libaries, eval=TRUE, include=FALSE}
library(car)
library(fBasics)
library(knitr)
library(car)
library(corrplot)
library(alr3)
library(MASS)
```


```{r,eval=TRUE, include=FALSE}

# read EVALUATION data set
eval_data <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-evaluation-data.csv")

# read training data set
mb_e <- read.csv("https://raw.githubusercontent.com/jtopor/CUNY-MSDA-621/master/HW-1/moneyball-training-data.csv")

#eliminate index column
# mb_e1 <- mb_e[,-1]
mb_e1 <- mb_e

#####Creating a new column for batting singles and eliminating hits for batting

#add singles column for hitting
mb_e1$TEAM_BATTING_1B <- as.numeric(mb_e1$TEAM_BATTING_H-mb_e1$TEAM_BATTING_2B-mb_e1$TEAM_BATTING_3B-mb_e1$TEAM_BATTING_HR)
mb_e1 <- mb_e1[,-3]
mb_e1 <- as.data.frame(mb_e1)

eval_data$TEAM_BATTING_1B <- as.numeric(eval_data$TEAM_BATTING_H - eval_data$TEAM_BATTING_2B - eval_data$TEAM_BATTING_3B - eval_data$TEAM_BATTING_HR)

# HITS is in second column in eval data
eval_data <- eval_data[,-2]

# ADD A DUMMY COLUMN TO EVAL DATA FOR TARGET WINS
eval_data$TARGET_WINS <- 0
```

#####Eliminate HBP, CS, and pitching HR's.

```{r,eval=TRUE,include=FALSE}
mb <- mb_e1[,-c(9,10,12)]
# summary(mb)

eval_data <- eval_data[,-c(8,9,11)]
# summary(eval_data)
```


#####Build model for batting SO using Gelman approach
```{r,eval=TRUE,include=FALSE}

#take out double plays + pitching SO + SB as data set is incomplete + Wins as they are not present in the evaluation data

BSO.1 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB -TARGET_WINS)
summary(BSO.1)

#eliminate doubles
BSO.2 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B)
summary(BSO.2)
vif(BSO.2)

# vif says remove TEAM_PITCHING_BB
BSO.3 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB)
summary(BSO.3)

# pvals say remove PITCHING_H
BSO.4 <- lm(data=mb, TEAM_BATTING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TARGET_WINS -TEAM_BATTING_2B - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(BSO.4)

vif(BSO.4)

##All p-values are low with a 686.8 F-statistic and adjusted R squared of 0.7236
#take a look
par(mfrow=c(2,2))
plot(BSO.2)

# ---------------------------------------
# function definition for impute function
impute <- function (a, a.impute){
  ifelse (is.na(a), a.impute,a)
}
# ---------------------------------------

#prediction function
pred.BSO <- round(predict(BSO.4, mb))
BSO.imp <- impute(mb$TEAM_BATTING_SO, pred.BSO)

# impute the evaluation data
pred_eval.BSO <- round(predict(BSO.4, eval_data))
eval.BSO.imp <- impute(eval_data$TEAM_BATTING_SO, pred_eval.BSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb$TEAM_BATTING_SO)
summary(BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb$TEAM_BATTING_SO, breaks = 200)
hist(BSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data$TEAM_BATTING_SO)
summary(eval.BSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data$TEAM_BATTING_SO, breaks = 30)
hist(eval.BSO.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb1 <- mb
mb1$TEAM_BATTING_SO <- BSO.imp

eval_data.1 <- eval_data
eval_data.1$TEAM_BATTING_SO <- eval.BSO.imp

```


#Build model for Pitching SO

```{r,eval=TRUE, include=FALSE}
#take out double plays + SB as data set is incomplete and wins as they are not present in evaluation data

PSO.1 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS)
summary(PSO.1)

vif(PSO.1)
# vif says remove TEAM_PITCHING_BB

PSO.2 <- lm(data=mb1, TEAM_PITCHING_SO~. - INDEX -TEAM_FIELDING_DP -TEAM_BASERUN_SB - TARGET_WINS - TEAM_PITCHING_BB)
summary(PSO.2)

vif(PSO.2)

#all low P value and F statistic of 4719 with adj R squared of 0.9952
#take a look
par(mfrow=c(2,2))
plot(PSO.2)

#place back in the data base with imputed data for SO's
pred.PSO <- round(predict(PSO.2, mb1))
PSO.imp <- impute(mb1$TEAM_PITCHING_SO, pred.PSO)

# impute the evaluation data
pred_eval.PSO <- round(predict(PSO.2, eval_data.1))
eval.PSO.imp <- impute(eval_data.1$TEAM_PITCHING_SO, pred_eval.PSO)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb1$TEAM_PITCHING_SO)
summary(PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb1$TEAM_PITCHING_SO, breaks = 200)
hist(PSO.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.1$TEAM_PITCHING_SO)
summary(eval.PSO.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.1$TEAM_PITCHING_SO, breaks = 30)
hist(eval.PSO.imp, breaks = 30)

###################################################

# update dataframes with imputed values 

mb2 <- mb1
mb2$TEAM_PITCHING_SO <- PSO.imp

eval_data.2 <- eval_data.1
eval_data.2$TEAM_PITCHING_SO <- eval.PSO.imp
```

#####Build model for SB
```{r,eval=TRUE,include=FALSE}
#Take out DP as incomplete data and target wins
SB.1 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP - TARGET_WINS)
summary(SB.1)

#eliminate pitching BB's
SB.2 <- lm(data=mb2, TEAM_BASERUN_SB~. -INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS)
summary(SB.2)

#eliminate singles
SB.3 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS)
summary(SB.3)

#simplify the model by taking out pitching
SB.4 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_1B - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.4)

#add singles back in
SB.5 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.5)

#eliminate doubles
SB.6 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B)
summary(SB.6)

#eliminate walks
SB.7 <- lm(data=mb2, TEAM_BASERUN_SB~. - INDEX -TEAM_FIELDING_DP -TEAM_PITCHING_BB - TARGET_WINS - TEAM_PITCHING_SO - TEAM_PITCHING_H - TEAM_BATTING_2B - TEAM_BATTING_BB)
summary(SB.7)

#all low P value and F statistic of 202.9 with adj R squared of 0.3427
#take a look
par(mfrow=c(2,2))
plot(SB.7)

#place back in the data base with imputed data for SB's
pred.SB <- round(predict(SB.7, mb2))
SB.imp <- impute(mb2$TEAM_BASERUN_SB, pred.SB)

# impute the evaluation data
pred_eval.SB <- round(predict(SB.7, eval_data.2))
eval.SB.imp <- impute(eval_data.2$TEAM_BASERUN_SB, pred_eval.SB)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb2$TEAM_BASERUN_SB)
summary(SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb2$TEAM_BASERUN_SB, breaks = 200)
hist(SB.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.2$TEAM_BASERUN_SB)
summary(eval.SB.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.2$TEAM_BASERUN_SB, breaks = 30)
hist(eval.SB.imp, breaks = 30)
###################################################

# update dataframes with imputed values
mb3 <- mb2
mb3$TEAM_BASERUN_SB <- SB.imp

eval_data.3 <- eval_data.2
eval_data.3$TEAM_BASERUN_SB <- eval.SB.imp
```

#####Build model to replace DP
```{r,eval=TRUE,include=FALSE}

#remove target wins
DP.1 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS)
summary(DP.1)

#remove batting 2B's
DP.2 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS - TEAM_BATTING_2B)
summary(DP.2)
# results show that EVERYTHING ELSE is statistically signficant, so:

# run vif to check for collinearity
vif(DP.2)
# results show TEAM_BATTING_SO should be removed

# remove TEAM_BATTING_SO
DP.3 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO)
summary(DP.3)
# p-value says remove TEAM_PITCHING_SO;


# remove TEAM_PITCHING_SO
DP.4 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TEAM_BATTING_2B -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO)
summary(DP.4)
vif(DP.4)
# P values and vif both indicate remove TEAM_PITCHING_BB

# remove TEAM_PITCHING_BB
DP.5 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(DP.5)
vif(DP.5)
# vif says remove TEAM_FIELDING_E; p-values all < .05 so remove TEAM_FIELDING_E

DP.6 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E)
summary(DP.6)
vif(DP.6)
# now no collinearity but p-values say remove TEAM_PITCHING_H

DP.7 <- lm(data=mb3, TEAM_FIELDING_DP~. - INDEX -TARGET_WINS -TEAM_BATTING_2B - TEAM_BATTING_SO - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_FIELDING_E - TEAM_PITCHING_H)
summary(DP.7)
vif(DP.7)
# no collinearity, all p-values < .05 so stop


#all low P value and F statistic of 255.8 with adj R squared of 0.3904
#take a look
par(mfrow=c(2,2))
plot(DP.7)


#place back in the data base with imputed data for SB's
# NOTE: Changed DP.4 to DP.7 here
pred.DP <- round(predict(DP.7, mb3))
DP.imp <- impute(mb3$TEAM_FIELDING_DP, pred.DP)

# impute the evaluation data
pred_eval.DP <- round(predict(DP.7, eval_data.3))
eval.DP.imp <- impute(eval_data.3$TEAM_FIELDING_DP, pred_eval.DP)

###################################################
# Jims added code for diagnostics of imputation

# first, check summaries to ensure similar values
summary(mb3$TEAM_FIELDING_DP)
summary(DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(mb3$TEAM_FIELDING_DP, breaks = 200)
hist(DP.imp, breaks = 200)

# ------------------ eval data checks ------------------------
# first, check summaries to ensure similar values
summary(eval_data.3$TEAM_FIELDING_DP)
summary(eval.DP.imp)

# now plot side-by-side histograms to check similarity of distributions
par(mfrow = c(2,2))
hist(eval_data.3$TEAM_FIELDING_DP, breaks = 30)
hist(eval.DP.imp, breaks = 30)
###################################################

# update data frames with imputed values
mb4 <- mb3
mb4$TEAM_FIELDING_DP <- DP.imp

eval_data.4 <- eval_data.3
eval_data.4$TEAM_FIELDING_DP <- eval.DP.imp
```


#####Eliminate unhistorical outliers - DO THIS FOR THE EVAL DATA AS WELL

```{r,eval=TRUE,include=FALSE}

# check rowcount before removal of outliers
nrow(mb4)
nrow(eval_data.4)

############## TEAM PITCHING_SO ############################
#most pitching SO's is 1450.  So delete all records with more than 1450 pitching SO's
mb5 <- mb4

# fixed error in this line: dataframe in 'which' call was mb1 so changed to mb5
mb5 <- mb5[which(mb5$TEAM_PITCHING_SO < 1450),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_SO < 1450),]

# check rowcount
nrow(mb5)
nrow(eval_data.4)

############ TEAM_PITCHING_H ##############################
#most ever hits by a team is 1730.  So delete all pitching hits >3000 to be conservative with the median
mb6 <- mb5
mb6 <- mb6[which(mb6$TEAM_PITCHING_H < 3001),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_PITCHING_H < 3001),]

# check rowcount
nrow(mb6)
nrow(eval_data.4)

############ TEAM_FIELDING_E ##############################
#most ever errors by a team is 639 by 1883 Philadelphia.  Prorating to 162 games gives a value of 1046.
mb7 <- mb6
mb7 <- mb7[which(mb7$TEAM_FIELDING_E < 1047),]

eval_data.4 <- eval_data.4[which(eval_data.4$TEAM_FIELDING_E < 1047),]

# ----------------------------------------------------------------------
# ----------------------------------------------------------------------

# check rowcount: result is 2172 => removed total of 104 rows
nrow(mb7)
nrow(eval_data.4)

dim(mb)-dim(mb7)

#we removed 104 rows total due to outliers in TRAINING data set.

# we removed 11 rows from the EVALUATION data set

# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(mb7) <- 1:nrow(mb7)
rownames(eval_data.4) <- 1:nrow(eval_data.4)

# drop INDEX column from training set
# mb7 <- mb7[,-1]

# now drop dummy column from evaluation data
# eval_data.4 <- eval_data.4[,-14]

# create CSV files containing updated data sets
write.csv(mb7, file = "/Users/scottkarr/IS621Summer2016/HW1/621-HW1-Clean-Data.csv", row.names = FALSE, col.names = TRUE)

write.csv(eval_data.4, file = "/Users/scottkarr/IS621Summer2016/HW1/621-HW1-Clean-EvalData-.csv", row.names = FALSE, col.names = TRUE)
```

# SINGLE PREDICTOR ANALYSIS & TRANSFORMATIONS
#####Model SMK Generlized Equation
Review descriptive statistics to confirm each variable is within acceptable bounds and 
contains no missing data.  Review Density plots of 13 variables for skewness to identify
which may require transformation.  

```{r full, eval = TRUE,include=FALSE, echo=FALSE}
#assign model to "clean" data set 
lm.smk <- mb7
#remove bad leverage points from diagnostic tests
lm.smk <- lm.smk[-c(1737,1920,226,391,385,1702,840,602,1928,1937,2109,2128,269,711,125),]
# now renumber rows of dataframe so that there are no gaps in row numbers
rownames(lm.smk) <- 1:nrow(lm.smk)
nrow(lm.smk)

#SINGLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.1B <- lm.smk$TEAM_BATTING_1B
m1 <- lm(mW~mb.1B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.1B),main="Singles");rug(mb.1B)
plot(mb.1B,StanRes1,xlab="Singles",ylab="Standardized Residuals");abline(lsfit(mb.1B,StanRes1),lty=2,col=2)
qqnorm(mb.1B,ylab="Y");qqline(mb.1B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.1B, family="bcPower")$lambda# round(-1.983968) => -2 => 1/(y^2)
tmb.1B <- mb.1B^(-2)
plot(density(tmb.1B),main="Singles");rug(tmb.1B)
plot(tmb.1B,StanRes1,xlab="Singles",ylab="Standardized Residuals");abline(lsfit(tmb.1B,StanRes1),lty=2,col=2)
qqnorm(tmb.1B,ylab="Y");qqline(tmb.1B,lty=2,col=2)

#DOUBLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.2B <- lm.smk$TEAM_BATTING_2B
m1 <- lm(mW~mb.2B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.2B),main="Doubles");rug(mb.2B)
plot(mb.2B,StanRes1,xlab="Doubles",ylab="Standardized Residuals");abline(lsfit(mb.2B,StanRes1),lty=2,col=2)
qqnorm(mb.2B,ylab="Y");qqline(mb.2B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.2B, family="bcPower")$lambda# round(0.5362315) => y^.5 => sqrt(y)
tmb.2B <- sqrt(mb.2B)
plot(density(tmb.2B),main="Doubles");rug(tmb.2B)
plot(tmb.2B,StanRes1,xlab="Doubles",ylab="Standardized Residuals");abline(lsfit(tmb.2B,StanRes1),lty=2,col=2)
qqnorm(tmb.2B,ylab="Y");qqline(tmb.2B,lty=2,col=2)

#TRIPLES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.3B <- lm.smk$TEAM_BATTING_3B
m1 <- lm(mW~mb.3B, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.3B),main="Triples");rug(mb.3B)
plot(mb.3B,StanRes1,xlab="Triples",ylab="Standardized Residuals");abline(lsfit(mb.3B,StanRes1),lty=2,col=2)
qqnorm(mb.3B,ylab="Y");qqline(mb.3B,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.3B, family="bcPower")$lambda# round(-0.03308475) => (1/y^30)
tmb.3B <- mb.3B^(-30)
plot(density(tmb.3B),main="Triples");rug(tmb.3B)
plot(tmb.3B,StanRes1,xlab="Triples",ylab="Standardized Residuals");abline(lsfit(tmb.3B,StanRes1),lty=2,col=2)
qqnorm(tmb.3B,ylab="Y");qqline(tmb.3B,lty=2,col=2)

#HOMERUNS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.HR <- lm.smk$TEAM_BATTING_HR
m1 <- lm(mW~mb.HR, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.HR),main="Homeruns");rug(mb.HR)
plot(mb.HR,StanRes1,xlab="Homeruns",ylab="Standardized Residuals");abline(lsfit(mb.HR,StanRes1),lty=2,col=2)
qqnorm(mb.HR,ylab="Y");qqline(mb.HR,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.HR, family="bcPower")$lambda# round(0.6348318) => y^(2/3)
tmb.HR <- mb.HR^(2/3)
plot(density(tmb.HR),main="Homeruns");rug(tmb.HR)
plot(tmb.HR,StanRes1,xlab="Homeruns",ylab="Standardized Residuals");abline(lsfit(tmb.HR,StanRes1),lty=2,col=2)
qqnorm(tmb.HR,ylab="Y");qqline(tmb.HR,lty=2,col=2)

#SLUGGING
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SL <- lm.smk$TEAM_BATTING_HR + 2 * lm.smk$TEAM_BATTING_3B 
m1 <- lm(mW~mb.SL, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SL),main="Slugging");rug(mb.SL)
plot(mb.SL,StanRes1,xlab="Slugging",ylab="Standardized Residuals");abline(lsfit(mb.SL,StanRes1),lty=2,col=2)
qqnorm(mb.SL,ylab="Y");qqline(mb.SL,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SL, family="bcPower")$lambda# round(0.5562253) => y^(.5)
tmb.SL <- sqrt(mb.SL)
plot(density(tmb.SL),main="Slugging");rug(tmb.SL)
plot(tmb.SL,StanRes1,xlab="Slugging",ylab="Standardized Residuals");abline(lsfit(tmb.SL,StanRes1),lty=2,col=2)
qqnorm(tmb.SL,ylab="Y");qqline(tmb.SL,lty=2,col=2)

#WALKS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.BB <- lm.smk$TEAM_BATTING_BB
m1 <- lm(mW~mb.BB, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.BB),main="Walks");rug(mb.BB)
plot(mb.BB,StanRes1,xlab="Walks",ylab="Standardized Residuals");abline(lsfit(mb.BB,StanRes1),lty=2,col=2)
qqnorm(mb.BB,ylab="Y");qqline(mb.BB,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.BB, family="bcPower")$lambda# round(1.434735) => y^(3/2)
tmb.BB <- mb.BB^(3/2)
plot(density(tmb.BB),main="Walks");rug(tmb.BB)
plot(tmb.BB,StanRes1,xlab="Walks",ylab="Standardized Residuals");abline(lsfit(tmb.BB,StanRes1),lty=2,col=2)
qqnorm(tmb.BB,ylab="Y");qqline(tmb.BB,lty=2,col=2)

#STRIKEOUTS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SO <- lm.smk$TEAM_BATTING_SO
m1 <- lm(mW~mb.SO, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SO),main="Strikeouts");rug(mb.SO)
plot(mb.SO,StanRes1,xlab="Strikeouts",ylab="Standardized Residuals");abline(lsfit(mb.SO,StanRes1),lty=2,col=2)
qqnorm(mb.SO,ylab="Y");qqline(mb.SO,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SO, family="bcPower")$lambda# round(0.7159533) => y^(3/4)
tmb.SO <- mb.SO^(3/4)
plot(density(tmb.SO),main="Strikeouts");rug(tmb.SO)
plot(tmb.SO,StanRes1,xlab="Strikeouts",ylab="Standardized Residuals");abline(lsfit(tmb.SO,StanRes1),lty=2,col=2)
qqnorm(tmb.SO,ylab="Y");qqline(tmb.SO,lty=2,col=2)

#STOLEN BASES:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mb.SB <- lm.smk$TEAM_BASERUN_SB
m1 <- lm(mW~mb.SB, data = lm.smk)
StanRes1 <- rstandard(m1)
plot(density(mb.SB),main="Stolen Bases");rug(mb.SB)
plot(mb.SB,StanRes1,xlab="Stolen Bases",ylab="Standardized Residuals");abline(lsfit(mb.SB,StanRes1),lty=2,col=2)
qqnorm(mb.SB,ylab="Y");qqline(mb.SB,lty=2,col=2)
##TRANSFORMATION
powerTransform(mb.SB, family="bcPower")$lambda# round(-0.03916886) => y^(-1/25)
tmb.SB <- mb.SB^(-1/25)
plot(density(tmb.SB),main="Stolen Bases");rug(tmb.SB)
plot(tmb.SB,StanRes1,xlab="Stolen Bases",ylab="Standardized Residuals");abline(lsfit(tmb.SB,StanRes1),lty=2,col=2)
qqnorm(tmb.SB,ylab="Y");qqline(tmb.SB,lty=2,col=2)

#PITCHING HITS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.H <- lm.smk$TEAM_PITCHING_H
m1 <- lm(mW~mp.H, data = lm.smk)
plot(density(mp.H),main="Pitching Hits");rug(mp.H)
plot(mp.H,StanRes1,xlab="Pitching Hits",ylab="Standardized Residuals");abline(lsfit(mp.H,StanRes1),lty=2,col=2)
qqnorm(mp.H, ylab = "Y");qqline(mp.H, lty = 2, col=2)
##TRANSFORMATION
powerTransform(mp.H,family="bcPower")$lambda# round(-3.097364) => y^(-3)
tmp.H <- mp.H^(-3)
plot(density(tmp.H),main="Pitching Hits");rug(tmp.H)
plot(tmp.H,StanRes1,xlab="Pitching Hits",ylab="Standardized Residuals");abline(lsfit(tmp.H,StanRes1),lty=2,col=2)
qqnorm(tmp.H,ylab="Y");qqline(tmp.H,lty=2,col=2)

#PITCHING WALKS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.BB <- lm.smk$TEAM_PITCHING_BB
m1 <- lm(mW~mp.BB, data = lm.smk)
plot(density(mp.BB),main="Pitching Walks");rug(mp.BB)
plot(mp.BB,StanRes1,xlab="Pitching Walks",ylab="Standardized Residuals");abline(lsfit(mp.BB,StanRes1),lty=2,col=2)
qqnorm(mp.BB, ylab = "Y");qqline(mp.BB,lty = 2,col=2)
##TRANSFORMATION
powerTransform(mp.BB, family="bcPower")$lambda# round(0.1609713) => y^(1/6)
tmp.BB <- mp.BB^(1/6)
plot(density(tmp.BB),main="Pitching Walks");rug(tmp.BB)
plot(tmp.BB,StanRes1,xlab="Pitching Walks",ylab="Standardized Residuals");abline(lsfit(tmp.BB,StanRes1),lty=2,col=2)
qqnorm(tmp.BB,ylab="Y");qqline(tmp.BB,lty=2,col=2)

#PITCHING STRIKEOUTS:
par(mfrow=c(2,3))
mW  <- lm.smk$TARGET_WINS
mp.SO <- lm.smk$TEAM_PITCHING_SO
m1 <- lm(mW~mp.SO, data = lm.smk)
plot(density(mp.SO),main="Pitching Strikeouts");rug(mp.SO)
plot(mp.SO,StanRes1,xlab="Pitching Strikeouts",ylab="Standardized Residuals");abline(lsfit(mp.SO,StanRes1),lty=2,col=2)
qqnorm(mp.SO, ylab = "Y");qqline(mp.SO,lty = 2,col=2)
##TRANSFORMATION
powerTransform(mp.SO, family="bcPower")$lambda#round(0.6522561) => y^(2/3)
tmp.SO <- mp.SO^(2/3)
plot(density(tmp.SO),main="Pitching Strikeouts");rug(tmp.SO)
plot(tmp.SO,StanRes1,xlab="Pitching Strikeouts",ylab="Standardized Residuals");abline(lsfit(tmp.SO,StanRes1),lty=2,col=2)
qqnorm(tmp.SO,ylab="Y");qqline(tmp.SO,lty=2,col=2)

#FIELDING ERRORS:
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.E <- lm.smk$TEAM_FIELDING_E 
m1 <- lm(mW~mf.E, data = lm.smk)
plot(density(mf.E),main="Fielding Errors");rug(mf.E)
plot(mf.E,StanRes1,xlab="Fielding Errors",ylab="Standardized Residuals");abline(lsfit(mf.E,StanRes1),lty=2,col=2)
qqnorm(mf.E, ylab = "Y");qqline(mf.E, lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.E,family="bcPower")$lambda#round(-0.939657) => (-9/10)
tmf.E <- mf.E^(-(9/10))
plot(density(tmf.E),main="Fielding Errors");rug(tmf.E)
plot(tmf.E,StanRes1,xlab="Fielding Errors",ylab="Standardized Residuals");abline(lsfit(tmf.E,StanRes1),lty=2,col=2)
qqnorm(tmf.E,ylab="Y");qqline(tmf.E,lty=2,col=2)

#DOUBLE PLAYS:
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.DP <- lm.smk$TEAM_FIELDING_DP
m1 <- lm(mW~mf.DP, data = lm.smk)
plot(density(mf.DP),main="Fielding Doubleplays");rug(mf.DP)
plot(mf.DP,StanRes1,xlab="Fielding Doubleplays",ylab="Standardized Residuals");abline(lsfit(mf.DP,StanRes1),lty=2,col=2)
qqnorm(mf.DP, ylab = "Y");qqline(mf.DP,lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.DP,family="bcPower")$lambda# round(1.49645) => 1.5
tmf.DP <- mf.DP^(3/2)
plot(density(tmf.DP),main="Fielding Doubleplays");rug(tmf.DP)
plot(tmf.DP,StanRes1,xlab="Fielding Doubleplays",ylab="Standardized Residuals");abline(lsfit(tmf.DP,StanRes1),lty=2,col=2)
qqnorm(tmf.DP,ylab="Y");qqline(tmf.DP,lty=2,col=2)

#FIELDING YIELD
par(mfrow=c(2,3))
mW   <- lm.smk$TARGET_WINS
mf.FY <- lm.smk$TEAM_FIELDING_E + lm.smk$TEAM_FIELDING_DP * 2
m1 <- lm(mW~mf.FY, data = lm.smk)
plot(density(mf.FY),main="Fielding Yield");rug(mf.FY)
plot(mf.FY,StanRes1,xlab="Fielding Yield",ylab="Standardized Residuals");abline(lsfit(mf.FY,StanRes1),lty=2,col=2)
qqnorm(mf.FY, ylab = "Y");qqline(mf.FY,lty = 2, col=2)
##TRANSFORMATION
powerTransform(mf.FY,family="bcPower")$lambda# round(-2.066982) => -2
tmf.FY <- mf.FY^(-2)
plot(density(tmf.FY),main="Fielding Yield");rug(tmf.FY)
plot(tmf.FY,StanRes1,xlab="Fielding Yield",ylab="Standardized Residuals");abline(lsfit(tmf.FY,StanRes1),lty=2,col=2)
qqnorm(tmf.FY,ylab="Y");qqline(tmf.FY,lty=2,col=2)

par(mfrow=c(1,1))
```

##### Evaluate Correlations  
Evaluate Correlation between predictors so as to not introduce collinearity into the model.
```{r correlation, eval=TRUE, echo=FALSE}
# assign model one "lm.1" to data set with all NA's imputed and "bad" leverage points removed
par(cex = 0.65)
corrplot(
    cor(lm.smk[c(1:13)]), 
    type='lower', 
    tl.srt=45,
    addshade="positive",
    addCoef.col = rgb(0,0,0,alpha=0.3),
    addCoefasPercent = TRUE)
nrow(lm.smk)
```

##### Model Selection Strategy
Two common strategies for adding or removing variables in a multiple regression model are
called backward elimination and forward selection. These techniques are often referred to
as stepwise model selection strategies, because they add or delete one variable at a time
as they “step” through the candidate predictors.  Model 1 uses the forward selection strategy 
which adds variables two-at-a-time until variables cannot be found that improve the model 
as measured by adjusted $R^2$.
Diez, D.M., Barr, C.D., & Çetinkaya-Rundel, M. (2015). OpenIntro Statistics (3rd Ed). pg. 378

#####Start with p.Hits & p.Walks
```{r m1-iteration1,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Hits <- tmp.H 
p.Walks <- tmp.BB
m1 <- lm(Wins ~ p.Hits+p.Walks)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Hits+p.Walks)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

##DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Hits,StanRest,ylab="Standardized Residuals")
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Add b.Singles & b.Doubles
```{r m1-interation2,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Hits <- tmp.H 
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
m1 <- lm(Wins ~ p.Hits+p.Walks+b.Singles+b.Doubles)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Hits+p.Walks+b.Singles+b.Doubles)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values of p.Hits >  0.05 so it gets removed
```

#####Removed p.Hits
```{r m1-interation3,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5 and adjusted $R^2$ increased

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
##Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Added Stolen Bases and Double Plays
```{r m1-interation4,eval=TRUE,include=FALSE}
nrow(lm.smk)
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)
#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
#If plot doesn't shows a linear relationship with no pattern or skew the model lacks normality.

# normal distribution, and uniform distribution of residuals
```

#####Added b.Walks and p.Strikeouts
```{r m1-interation5,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
b.Walks <- tmb.BB
p.StrikeOuts <- tmp.SO
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.Walks+p.StrikeOuts)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.Walks+p.StrikeOuts)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 but VIFs  > 5
#highest vif is b.Walks so it gets removed
```

#####Remove b.Walks
```{r m1-interation6,eval=TRUE,include=FALSE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB 
f.DoublePlays <- tmf.DP
p.StrikeOuts <- tmp.SO
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+p.StrikeOuts)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+p.StrikeOuts)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-value for b.Walks > 0.05 all VIFs < 5
```

#####Add b.StrikeOuts, & b.Slugging
```{r m1-interation7,eval=TRUE,include=FALSE}
nrow(lm.smk)
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
f.DoublePlays <- tmf.DP
b.StrikeOuts <- tmb.SO
b.Slugging <- tmb.SL
m1 <- lm(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+f.DoublePlays+b.StrikeOuts+b.Slugging)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.Doubles+b.StolenBases+b.StrikeOuts+b.Slugging)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)

#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(b.StrikeOuts,StanRest,ylab="Standardized Residuals")
plot(b.Slugging,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

#####Add b.Fielding
\[
\begin{aligned}
\widehat{wins} &= \hat{\beta}_0 + 
                    \hat{\beta}_1 \times p.Walks 
                    \hat{\beta}_2 \times b.Singles + 
                    \hat{\beta}_3 \times b.Doubles +
                    \hat{\beta}_4 \times b.Stolen Bases +                    
                    \end{aligned}
\]
\[
\begin{aligned}
                    \hat{\beta}_5 \times f.Double Plays +  
                    \hat{\beta}_6 \times b.Strike Outs +
                    \hat{\beta}_7 \times b.Slugging +   
                    \hat{\beta}_8 \times b.Fielding Yield +                       
                    \end{aligned}
\]
nrow(lm.smk)
```{r m1-interation8,eval=TRUE}
#VARIABLES
#variables have been transformed first as individual predictors
Wins  <- mW
Index <- lm.smk$INDEX 
p.Walks <- tmp.BB
b.Singles <- tmb.1B
b.Doubles <- tmb.2B
b.StolenBases <- tmb.SB
b.StrikeOuts <- tmb.SO
b.Slugging <- tmb.SL
f.Fielding <- tmf.FY

m1 <- lm(Wins ~ -Index+p.Walks+b.Singles+b.Doubles+b.StolenBases+b.StrikeOuts+b.Slugging+f.Fielding)

#PAIRWISE PLOT
par(mfrow=c(1,1))
pairs(Wins ~ p.Walks+b.Singles+b.StolenBases+b.StrikeOuts+b.Slugging+f.Fielding)

#MODEL DIAGNOSTICS
summary(m1)

#DIAGNOSTIC1. show collinearity of variables after checking p-values to < 0.05.
vif(m1)
#p-values are all < 0.05 and no VIFs  > 5

#DIAGNOSTIC2.  generate Added Variable Plots: should show linear relationship between response & predictors:
par(mfrow=c(2,2))
avPlots(m1, ~.,ask=FALSE, id.n = 2)

#relationship is linear

#DIAGNOSTIC3.  generate Summary Diagnostic Plots
par(mfrow=c(2,2))
plot(m1)
#Upper Left plot "Residuals vs Fitted" 
# clear predictable pattern
# uniform variability for all fitted values
#Upper Right 
#  normality in residuals
#Lower Right plot "Residuals vs. Leverage"
#  normal distribution, and uniform distribution of residuals
#  no significant leverage points

#DIAGNOSTIC4.  generate Standardized Residual Plots against each predictor
par(mfrow=c(2,2))
StanRest <- rstandard(m1)
plot(p.Walks,StanRest,ylab="Standardized Residuals")
plot(b.Singles,StanRest,ylab="Standardized Residuals")
plot(b.Doubles,StanRest,ylab="Standardized Residuals")
plot(b.StolenBases,StanRest,ylab="Standardized Residuals")
plot(f.DoublePlays,StanRest,ylab="Standardized Residuals")
plot(b.StrikeOuts,StanRest,ylab="Standardized Residuals")
plot(b.Slugging,StanRest,ylab="Standardized Residuals")
plot(f.Fielding,StanRest,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRest,ylab="Standardized Residuals",xlab="Fitted Values")
#Examine plots for constant variability of residuals across ALL predictor. 
# uniform distribution of residuals

#DIAGNOSTIC5.  generate plot of Y "response variable"" against Fitted Values "regression model"
par(mfrow = c(2,2))
plot(m1$fitted.values,Wins,xlab="Fitted Values",ylab=expression(Wins^lambda))
abline(lsfit(m1$fitted.values,Wins))
plot(m1)
# normal distribution, and uniform distribution of residuals
```

```{r}
#pred_eval.m1 <- round(predict(m1, eval_data))
#eval.BSO.imp <- impute(eval_data$TARGET_WINS, pred_eval.m1)
```