---
title: "JT HW1 ALL MODELS"
author: "James Topor"
date: "June 17, 2016"
output: html_document
---

# --------------------------------------------------------------------------------------------

# Model # 1: General Model - All Variables


```{r}
library(car)

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```

```{r, eval = FALSE, echo = FALSE}

library(alr3)

# code for finding power transforms for skewed variables
hist(mb_clean$TEAM_PITCHING_BB, breaks = 200)

hist(mb_clean$TEAM_PITCHING_H, breaks = 200)

hist(mb_clean$TEAM_FIELDING_E, breaks = 200)

############### TEAM_PITCHING_BB ##############################

summary( powerTransform( cbind(TEAM_PITCHING_BB) ~ 1, mb_clean))

# TEAM_PITCHING_BB = log transform
PBB.T <- log(mb_clean$TEAM_PITCHING_BB)
hist(PBB.T, breaks = 200)

# ---------------------
############### PITCHING_H ###################################
summary( powerTransform( cbind(TEAM_PITCHING_H) ~ 1, mb_clean))
# TEAM_PITCHING_H = 1/y^3

# WORKS! 
PH.T <- 1/(mb_clean$TEAM_PITCHING_H ^ 3)
hist(PH.T, breaks = 200)


############## FIELDING_E #####################################
# -------------------------
summary( powerTransform( cbind(TEAM_FIELDING_E) ~ 1, mb_clean))
# TEAM_FIELDING_E = -1 = 1/y or try log

# WORKS!
FE.T <- 1/mb_clean$TEAM_FIELDING_E
hist(FE.T, breaks = 200)

# Now load transformed values into data set

mb.t <- mb_clean

mb.t$TEAM_PITCHING_BB <- PBB.T
mb.t$TEAM_PITCHING_H <- PH.T
mb.t$TEAM_FIELDING_E <- FE.T
```
# ----------------------------------------------------------------------------------------
# ----------------------------------------------------------------------------------------

## Build Models

### Model using all remaining variables as a starting point

__Yields r^2= 0.3347, Adj r^2 = 0.3322, F = 136__

```{r}

# keep the clean data set pure
mb <- mb_clean

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# p-value indicates remove TEAM_BATTING_1B

# --------------------
# remove TEAM_BATTING_1B
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_BATTING_1B)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_SO or PITCHING_SO, so remove PITCHING_SO per other models

# -------------------
#eliminate TEAM_PITCHING_SO
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO)
summary(model.3)

vif(model.3)

# vif says remove TEAM_BATTING_BB or PITCHINNG_BB so go with PITCHING_BB

# -------------------
#eliminate TEAM_PITCHING_BB
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# --------------------
# eliminate TEAM_BATTING_2B
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_BATTING_2B)
summary(model.5)

# p-values < .05 so check for collinearity
vif(model.5)

# no colinearity so STOP HERE
```


## DIAGNOSTICS

Plots are linear except for errors
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.5, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS
Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. Outliers at 1920, 1737, 393, 1515
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.5)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, SB, Pitch_H, Fielding_E, BATT_BB, BATT_SO, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.5)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.5$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```

Now cleanup data objects that are no longer required
```{r}
rm(model, model.2, model.3, model.4)
```


# -----------------------------------------------------------------------------------------

## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 1920, 1737, 393, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_clean[-c(1920, 1737, 393, 1515),]

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```

# -----------------------------------------------------------------------------------

## Now refit first model from above: all variables

__Yields r^2= 0.3504, Adj r^2 = 0.348, F = 145.6__

```{r}

# keep the clean data set pure
mb <- mb_rem

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
vif(model)
# vif indicates remove TEAM_PITCHING_SO

# --------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# p-values say remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)
vif(model.4)

# pvals say remove TEAM_BATTING_2B
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_BATTING_2B - TEAM_BATTING_1B - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.5)
vif(model.5)


# vif and pvals OK so stop
```


## SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. __Might not be a good model__.
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.4)
```

# ------------------------------------------------------------
## Plots show outliers so remove them and re-fit

Per Cooks Distance, remove items 1931, 391, 820, 1933, 835, 2124

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem2 <- mb[-c(1931, 391, 820, 1933, 835, 2124),]

# renumber rows
rownames(mb_rem2) <- 1:nrow(mb_rem2)
```

## Model using all remaining variables as a starting point

__Yields r^2= 0.3598, Adj r^2 = 0.3572, F = 134.4__

```{r}

# keep the clean data set pure
mb <- mb_rem2

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX )
summary(model)

# p-values all < .05 so check collinearity
vif(model)

# vif says remove TEAM_PITCHING_SO

# -------------------
#eliminate TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values OK so check collinearity
vif(model.2)
# vif says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_BATTING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# p-values says remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB - TEAM_BATTING_1B)
summary(model.4)

# p-values OK so check collinearity
vif(model.4)
# vif OK so STOP

# get MSE of residuals
anova(model.4)

```

## Diagnostics

Plots are linear except for errors
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Residuals appear to be within 2 std devs. 
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, SB, Pitch_H, Fielding_E, PITCH_BB, PITCH_SO, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_2B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```


# -------------------------------------------------------------------------------


## Now try same model but with FIELD_E transformed by Box-Cox recommended power transform

```{r}
# TEAM_FIELDING_E: Box-cox yields -1 => 1/y

mb$TEAM_FIELDING_E <- 1/mb$TEAM_FIELDING_E
```


__Yields r^2= 0.3168, Adj r^2 = 0.3143, F = 124.8__

```{r}

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
# p-values all < .05 so check collinearity

vif(model)

# vif says remove TEAM_PITCHING_SO

# -------------------
#eliminate TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values says remove TEAM_BATTING_1B

# -------------------
#eliminate TEAM_BATTING_1B
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B)
summary(model.3)
vif(model.3)

# p-values says remove TEAM_PITCHING_BB

# -------------------
#eliminate TEAM_BATTING_BB
model.4 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B - TEAM_PITCHING_BB)
summary(model.4)

# p-values say remove doubles
model.5 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_BATTING_1B - TEAM_PITCHING_BB - TEAM_BATTING_2B)
summary(model.5)

vif(model.5)

# vif OK so STOP

# turn off scientific formatting of results
options(scipen=999)
model.5
anova(model.5)
```

## Diagnostics

Plots are all linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.5, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in lower left end but not drastic; Residuals not all within 2 std devs.  
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.5)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for several variables: 3B, HR, BB, SO, SB, Pitch_H, Fielding_E
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.5)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_3B, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_HR, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")

```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship whose slope might be slightly skewed by outliers in upper right of plot
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.5$fitted.values
# nrow(fit1)

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```


```{r}
# clean up objects in memory
rm(list = ls())
```









# ------------------------------------------------------------------------------------------------
# ------------------------------------------------------------------------------------------------








# Model # 2: Total Bases






```{r}
library(car)

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```


# ----------------------------------------------------------------------------------------



## Build a model with Total Bases added and all of the other hitting vars removed

First, create the new variable and discard its components

```{r}
mb_t <- mb_clean

mb_t$TOTAL_BASES <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR)

# plot histogram to check shape of distribution
par(mfrow = c(1,1))
hist(mb_t$TOTAL_BASES, breaks = 200)

# now drop 1B, 2B, 3B, HR
mb_tb <- mb_t[,c(1, 2, 6, 7, 8, 9, 10, 11, 12, 13, 15)]

###################################################################
# check correlation with WINS and run simple linear model
cor(mb_tb$TARGET_WINS, mb_tb$TOTAL_BASES) 
mtest <- lm(data=mb_tb, TARGET_WINS ~ TOTAL_BASES)
summary(mtest) 

# shows .381 correlation and Adj R^2 of 0.1453

plot(mb_tb$TARGET_WINS ~ mb_tb$TOTAL_BASES)
abline(lm(mb_tb$TARGET_WINS ~ mb_tb$TOTAL_BASES), lty=2)
# plot doesn't show unusual relationship
######################################################################
```

__Yields r^2= 0.3175, Adj. R^2 = 0.3153, F = 143.8__

```{r}
# fit model
model <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX)
summary(model)

# All p-values < .05 so check collinearity
vif(model)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO. CHoose PITCHING_SO as in other models

# ----------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.3 <- lm(data=mb_tb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

# All p-values < .05 so check collinearity
vif(model.3)
# no further collinearity issues so STOP

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E  shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots look good except for outliers. Lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals except for outliers, most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.3)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for BASERUN_SB, PITCHING_H, PITCHING_SO, FIELDING_E, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.3)
par(mfrow=c(2,2))

plot(mb_tb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")

plot(mb_tb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Might be some skew due to outliers
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.3$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tb$TARGET_WINS),lty=2)
```


# -----------------------------------------------------------------------------------------
# -----------------------------------------------------------------------------------------

## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 1920, 1737, 393, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_tb[-c(1920, 1737, 393, 1515),]

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```

# -----------------------------------------------------------------------------------

## Now refit first model from above: all variables


__Yields r^2= 0.3287, Adj r^2 = 0.3265, F = 151.1__

```{r}

# keep the clean data set pure
mb <- mb_rem

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# pvals all < .05 so check collinearity

vif(model)
# vif indicates remove TEAM_BATTING_SO and PITCHING_SO. CHoose PITCHING_SO again

# --------------------
# remove TEAM_PITCHING_SO
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values are OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_BB

# -------------------
#eliminate TEAM_BATTING_BB or PITCHING_BB so choose PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_SO)
summary(model.3)

# p-values OK so check collinearity
vif(model.3)
# vif and pvals OK so STOP HERE

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted. Normal QQ shows a bit of skew in upper right end but not drastic; Some residuals appear to be outside 2 std devs. __Might not be a good model__. Outliers at 1528, 1922, 820, 1933, 1733, 835
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.3)
```


### Plots show outliers so remove them and re-fit

Per Cooks Distance, remove 1528, 1922, 820, 1933, 1733, 835

```{r}
############ SECOND SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem2 <- mb[-c(1528, 1922, 820, 1933, 1733, 835),]

# renumber rows
rownames(mb_rem2) <- 1:nrow(mb_rem2)
```

### Model using all remaining variables as a starting point

__Yields r^2= 0.33655, Adj r^2 = 0.3343, F = 156__

```{r}

# keep the clean data set pure
mb <- mb_rem2

# use p-value elimination
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)
# p-values all < .05 so check collinearity

vif(model)

# vif says remove TEAM_BATTING_SO

# -------------------
#eliminate TEAM_BATTING_SO or PITCHING_SO so choose PITCHING_SO again
model.2 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO)
summary(model.2)

# p-values OK so check collinearity
vif(model.2)
# vif says remove TEAM_BATTING_BB or PITCHING_BB so choose PITCHING_BB again

# -------------------
#eliminate TEAM_PITCHING_BB
model.3 <- lm(data=mb, TARGET_WINS ~. - INDEX - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.3)

vif(model.3)
# p-values and VIF OK so STOP

anova(model.3)

# check 95% confidence intervals for coefficients
confint(model.3)
```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.3, id.n = 2)
```

### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Lack of constant variability in Resid vs. Fitted but only for extreme outliers. Normal QQ looks very good; Some Residuals appear to be outside 2 std devs
```{r}
# plot summary residual plots
par(mfrow=c(2,2))
plot(model.3)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for BASERUN_SB, PITCHING_H, PITCHING_SO, FIELDING_E, FIELDING_DP
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.3)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_PITCHING_H, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Might be some skew due to outliers
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.3$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```
# -------------------------------------------------------------------------------------------
# -------------------------------------------------------------------------------------------

## Now try same model but with FIELD_E transformed according to Box-Cox

```{r}
# TEAM_FIELDING_E: Box-Cox yields power xform of -1 => 1/y

mb$TEAM_FIELDING_E <- 1/mb$TEAM_FIELDING_E
```

Now refit first model from above: all variables

## Model using all remaining variables as a starting point

__Yields r^2= 0.3048, Adj r^2 = 0.3029, F = 157.5__

```{r}

# fit model
model <- lm(data=mb, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

vif(model.2)
# vif says remove TEAM_BATTING_SO or PITCHING_SO so discard PITCHING_SO again

# ----------------------
# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

vif(model.3)
# vif say remove TEAM_BATTING_BB or PITCHING_BB - go with PITCHING_BB again

# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

options(scipen=999)
model.4
anova(model.4)

# no collinearity so STOP
# check 95% confidence intervals for coefficients
confint(model.4)

```

## Diagnostics

Plots show all variables are linear to response
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at both ends; normal distribution of residuals; many residuals outside of 2 std devs __PROBABLY NOT A GOOD MODEL__
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb$TEAM_BATTING_BB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_BASERUN_SB, StanRes1, ylab="Standardized Residuals")
plot(mb$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")

plot(mb$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb$TOTAL_BASES, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb$TARGET_WINS),lty=2)
```

```{r}
# clean up objects in memory
rm(list = ls())
```









# ------------------------------------------------------------------------------------------
# ------------------------------------------------------------------------------------------








# Model 3: Total Bases PLUS

```{r}
library(car)

# read clean data set from Github

mb_clean <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/621-HW1-Clean-Data.csv")  
```

# ---------------------------------------------------------------------------------------

## Build a model with Total Bases + SB + BB added and all of the other hitting vars removed

```{r}

# create new variable and drop its components
mb_t <- mb_clean

mb_t$TB_PLUS <- mb_clean$TEAM_BATTING_1B + (2 * mb_clean$TEAM_BATTING_2B) + 
                       (3 * mb_clean$TEAM_BATTING_3B) + (4 * mb_clean$TEAM_BATTING_HR) + 
                       mb_clean$TEAM_BATTING_BB + mb_clean$TEAM_BASERUN_SB

par(mfrow = c(1,1))
hist(mb_t$TB_PLUS, breaks = 200)

# now drop 1B, 2B, 3B, HR, BB, SB
mb_tbp <- mb_t[,c(1, 2, 7, 9, 10, 11, 12, 13, 15)]

###################################################################
# check correlation with WINS and run simple linear model
cor(mb_tbp$TARGET_WINS, mb_tbp$TB_PLUS) 
mtest <- lm(data=mb_tbp, TARGET_WINS ~ TB_PLUS)
summary(mtest) 

# shows .448 correlation and Adj R^2 of 0.2002 => better than TOTAL_BASES

plot(mb_tbp$TARGET_WINS ~ mb_tbp$TB_PLUS)
abline(lm(mb_tbp$TARGET_WINS ~ mb_tbp$TB_PLUS), lty=2)
# plot doesn't show unusual relationship
######################################################################
```

# -----------------------------------------------------------------------------

__Yields r^2= 0.2845, Adj. R^2 = 0.2832, F = 215.4__

```{r}
# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO, so try removing PITCHING_SO

# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

# p-vals say remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

# no collinearity so STOP

```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: outliers at 2012, 1820, 859. Lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals except for outliers, most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```


# -----------------------------------------------------------------------------------------
# -----------------------------------------------------------------------------------------

## REMOVE OUTLIERS AND REFIT

Per Cooks Distance, remove items 836, 821, 1920, 1737, 1515

```{r}
############ FIRST SET OF OUTLIERS ######################
# drop outlier records from data set
mb_rem <- mb_tbp[-c(836, 821, 1920, 1737, 1515),]

# save first data set
mb_tbp_orig <- mb_tbp

# renumber rows
rownames(mb_rem) <- 1:nrow(mb_rem)
```

# -----------------------------------------------------------------------------------

## Now refit first model from above: all variables

__Yields r^2= 0.2944, Adj r^2 = 0.2931, F = 225.5__

```{r}

# keep the clean data set pure
mb_tbp <- mb_rem

# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ .- INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H)
summary(model.2)

# All p-values < .05 so check collinearity
vif(model.2)
# vif indicates remove TEAM_BATTING_SO or TEAM_PITCHING_SO, so remove PITCHING_SO again

# ----------------------
# remove TEAM_PITCHING_SO
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.3)

# p-vals say remove TEAM_PITCHING_BB


# ----------------------
# remove TEAM_PITCHING_BB
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_H - TEAM_PITCHING_SO - TEAM_PITCHING_BB)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

anova(model)

# no collinearity so STOP

```

## Diagnostics

Plots for Fielding_E shows skew. Others are pretty linear
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals; most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_BATTING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```


# -------------------------------------------------------------------------------------------
# -------------------------------------------------------------------------------------------

## Now try same model but with FIELD_E transformed using Box-Cox

```{r}
# TEAM_FIELDING_E: Box-cox says -1 power transform => 1/y 

mb_tbp$TEAM_FIELDING_E <- 1/mb_tbp$TEAM_FIELDING_E
```


Now refit first model from above: Start with all variables

__Yields r^2= 0.2932, Adj r^2 = 0.2919, F = 224.3__

```{r}

# fit model
model <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX)
summary(model)

# p-vals say remove TEAM_PITCHING_BB

# ----------------------
# remove TEAM_PITCHING_BB
model.2 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB)
summary(model.2)

# p-values say remove TEAM_PITCHING_H

# ----------------------
# remove TEAM_PITCHING_H
model.3 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_H)
summary(model.3)

# p-vals say remove TEAM_PITCHING_SO

# ----------------------
# remove TEAM_PITCHING_SO
model.4 <- lm(data=mb_tbp, TARGET_WINS ~ . - INDEX - TEAM_PITCHING_BB - TEAM_PITCHING_H - TEAM_PITCHING_SO)
summary(model.4)

# pvals all < .05 so check collinearity
vif(model.4)

# no collinearity so STOP
options(scipen=999)
model.4
anova(model.4)
```

## Diagnostics

Plots show all variables are linear to response
```{r}
# CREATE ADDED VARIABLE PLOTS TO ASSESS predictor vs response
avPlots(model.4, id.n = 2)
```


### SUMMARY MODEL DIAGNOSTIC PLOTS

Plots: Some lack of Constant variability in Resid vs. Fitted at very large values of Yhat; normal distribution of residuals; most residuals within 2 std dev and well within Cook's distance
```{r}
#Figure 5.6 on page 129 MARR text
par(mfrow=c(2,2))
plot(model.4)
```

### PLOT STANDARDIZED RESIDUALS AGAINST EACH PREDICTOR

Results show lack of constant variability for PITCHING_SO, FIELDING_E, FIELDING_DP, TB_PLUS
```{r}

# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

StanRes1 <- rstandard(model.4)
par(mfrow=c(2,2))

plot(mb_tbp$TEAM_PITCHING_SO, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_E, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TEAM_FIELDING_DP, StanRes1, ylab="Standardized Residuals")
plot(mb_tbp$TB_PLUS, StanRes1, ylab="Standardized Residuals")
```

### PLOT Y AGAINST FITTED VALUES

Plot shows a linear relationship with no pattern or skew
```{r}
# get rows have no NA's from data frame
# NoNA <- mb_mods[!rowSums(is.na(mb_mods[1:13])), ]

fit1 <- model.4$fitted.values

par(mfrow = c(1,1))
plot(fit1, mb_tbp$TARGET_WINS,xlab="Fitted Values")
abline(lsfit(fit1, mb_tbp$TARGET_WINS),lty=2)
```

```{r}
# clean up objects in memory
rm(list = ls())
```

