---
title: "Moneyball Multiple Regression Model"
output:
  html_document:
    highlight: pygments
    keep_md: yes
    
    theme: cerulean
    toc: yes
  pdf_document: default
  word_document: default
---

```{r load-libaries, eval=TRUE, include=TRUE}
library(fBasics)
library(knitr)
library(car)
library(corrplot)
library(alr3)
```

Source files can be found at
[!github-link](https://github.com/scottkarr/IS621-HW1)

```{r load-data, eval=TRUE,include=TRUE}

BB.list <- read.csv("https://raw.githubusercontent.com/spsstudent15/2016-02-621-W1/master/moneyball-training-data.csv", header=TRUE,
    stringsAsFactors=FALSE, sep=",")
BB.df0 <- data.frame(BB.list)
#remove INDEX column
BB.df0 <- BB.df0[,-1]
BB.df0$TEAM_BATTING_1B <- # replace hits hits column with singles
  as.numeric(
      BB.df0$TEAM_BATTING_H-BB.df0$TEAM_BATTING_2B-BB.df0$TEAM_BATTING_3B-BB.df0$TEAM_BATTING_HR
  )
#remove TEAM_BATTING_H column
BB.df0 <- BB.df0[,c(-2)]
#remove TEAM_BASERUN_CS, TEAM_BATTING_HBP & TEAM_PITCHING_HR columns
BB.df0  <- BB.df0[,-c(8,9,11)]
BB.df <- BB.df0
DescStats.df <- basicStats(BB.df)
```

# Data Dictionary
variable            | definition                | effect      | Mo | M1 | M2
----------------    | ------------------------  | ----------- | -- | -- | --
`INDEX`             | Identification Variable   | None        | N  | N  | N
`TARGET_WINS`       | Number of wins            | Positive    | Y  | Y  | Y
`TEAM_BATTING_H`    | Base Hits by batters      | Removed     | N  | N  | N
`TEAM_BATTING_1B`   | Singles by batters (1B)   | Positive    | Y  | Y  | Y
`TEAM_BATTING_2B`   | Doubles by batters (2B)   | Positive    | Y  | Y  | Y
`TEAM_BATTING_3B`   | Triples by batters (3B)   | Positive    | Y  | Y  | Y
`TEAM_BATTING_HR`   | Homeruns by batters (4B)  | Positive    | Y  | Y  | Y
`TEAM_BATTING_BB`   | Walks by batters          | Positive    | Y  | Y  | Y
`TEAM_BATTING_SO`   | Strikeouts by batters     | Negative    | Y  | Y  | Y
`TEAM_BATTING_HBP`  | Batters hit by pitch      | Removed     | N  | N  | N
`TEAM_BASERUN_SB`   | Stolen bases              | Removed     | N  | N  | N
`TEAM_BASERUN_CS`   | Caught stealing           | Removed     | N  | N  | N
`TEAM_PITCHING_H`   | Hits allowed              | Negative    | Y  | Y  | Y
`TEAM_PITCHING_HR`  | Homeruns allowed          | Removed     | N  | N  | N
`TEAM_PITCHING_BB`  | Walks allowed             | Negative    | Y  | Y  | Y
`TEAM_PITCHING_SO`  | Strikeouts by pitchers    | Positive    | Y  | Y  | Y
`TEAM_FIELDING_E`   | Errors                    | Negative    | Y  | Y  | Y
`TEAM_FIELDING_DP`  | Double Plays              | Positive    | Y  | Y  | Y

# Exploring the data
Many of the statistics in the data sets provided have been extrapolated using base 
statistics from the deadball era circa ~1900-1920 and prior.  Outliers that need 
to be adjusted can be found using the reference link [!baseball-almanac](http://www.baseball-almanac.com/recbooks/rb_stba2.shtml).  Note 
that during the deadball era, a nearly soft ball was used which had dramatic effect on 
power hitting and pitching statistics.  Any adjustment that rationalize the data from 
this period into observations that include post WWII statistics should be bound by the 
later era's limits so distributions aren't skewed.


# Handle Missing Data
variable            | NA count         |  NA %       |  action
----------------    | ---------------- | ----------- | ---------------
`TEAM_BATTING_SO`   |  102             |  4.48       | impute w/ median
`TEAM_BASERUN_SB`   |  131             |  5.75       | impute w/ median
`TEAM_BASERUN_CS`   |  772             | 33.89       | removed variable
`TEAM_BATTING_HBP`  | 2085             | 91.53       | removed variable
`TEAM_PITCHING_SO`  |  102             |  4.48       | impute w/ median
`TEAM_FIELDING_DP`  |  286             | 12.55       | impute w/ median

Deleting missing cases is the simplest strategy for dealing with missing data.  It 
avoids the complexity and possible biases introduced by more sophisticated methods. 
The drawback is throwing away infomration that might allow more precise inference. 
If relatively few cases contain missing values deleting still leaves a large dataset 
or to communicate a simple data analysis method, the deltion strategy is satisfactory.  

Standard errors are larger after deleting cases because of fewer records to fit the model. 
Larger standard errors results in less precise estimates.  (Faraway, LMR 2015, p.200) 

Single imputation  . .  causes bias, while deletion causes a loss of information. 
Multiple imputation is a way to reduce the bias caused by single imputation.  The 
problem with single imputation is the value tends to be less variable than the value 
we would have seen because it does not include the error variation normally seen in 
observed data.  The idea behind multiple imputation is to reinclude that error variation. 
(Faraway, LMR 2015, p.202)  

Multiple imputation can be done using the Amelia package.  Per Faraway, the assumption is 
the data is multivariate normal, so heavily skewed varibales should be log-transformed first.  

#Impute Missing Data
#####Build model for batting SO using Gelman approach
```{r}
BSO <- lm(data=BB.df, TEAM_BATTING_SO~.)
summary(BSO)

#take out double plays + pitching SO + SB as data set is incomplete

BSO.1 <- lm(data=BB.df, TEAM_BATTING_SO~. -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB)
summary(BSO.1)

#eliminate doubles
BSO.2 <- lm(data=BB.df, TEAM_BATTING_SO~. -TEAM_FIELDING_DP -TEAM_PITCHING_SO -TEAM_BASERUN_SB - TEAM_BATTING_2B)
summary(BSO.2)

##All p-values are low with a 604.1 F-statistic and adjusted R squared of 0.7246
#take a look
par(mfrow=c(2,2))
plot(BSO.2)
plot(BSO.2$residuals)

#prediction function
pred.BSO <- round(predict(BSO.2, BB.df))
impute <- function (a, a.impute){
  ifelse (is.na(a), a.impute,a)
}

BSO.imp.1 <- impute(BB.df$TEAM_BATTING_SO, pred.BSO)

#place back in the data base with imputed data for SO's

BB.df$TEAM_BATTING_SO <- BSO.imp.1


```

#Build model for Pitching SO
```{r}
PSO <- lm(data=BB.df, TEAM_PITCHING_SO~.)
summary(PSO)

#take out double plays + SB as data set is incomplete

PSO.1 <- lm(data=BB.df, TEAM_PITCHING_SO~. -TEAM_FIELDING_DP -TEAM_BASERUN_SB)
summary(PSO.1)

#all low P value and F statistic of 43710 with adj R squared of 0.9958
#take a look
par(mfrow=c(2,2))
plot(PSO.1)
plot(PSO.1$residuals)


#place back in the data base with imputed data for SO's
pred.PSO <- round(predict(PSO.1, BB.df))
PSO.imp.1 <- impute(BB.df$TEAM_PITCHING_SO, pred.PSO)
BB.df$TEAM_PITCHING_SO <- PSO.imp.1
```

#####Build model for SB
```{r}
SB <- lm(data=BB.df, TEAM_BASERUN_SB~.)
summary(SB)

#Take out DP as incomplete data
SB.1 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP)
summary(SB.1)

#eliminate pitching BB's
SB.2 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB)
summary(SB.2)

#eliminate triples
SB.3 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B)
summary(SB.3)

#eliminate batting walks
SB.4 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB)
summary(SB.4)

#eliminate singles
SB.5 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB - TEAM_BATTING_1B)
summary(SB.5)

#eliminate doubles
SB.6 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB - TEAM_BATTING_1B - TEAM_BATTING_2B)
summary(SB.6)

#eliminate pitching SO
SB.7 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB - TEAM_BATTING_1B - TEAM_BATTING_2B - TEAM_PITCHING_SO)
summary(SB.7)

#add singles back in
SB.8 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB - TEAM_BATTING_2B - TEAM_PITCHING_SO)
summary(SB.8)

#take pitching hits out
SB.9 <- lm(data=BB.df, TEAM_BASERUN_SB~.-TEAM_FIELDING_DP -TEAM_PITCHING_BB -TEAM_BATTING_3B -TEAM_BATTING_BB - TEAM_BATTING_2B - TEAM_PITCHING_SO - TEAM_PITCHING_H)
summary(SB.9)

#all low P value and F statistic of 303.5 with adj R squared of 0.4386
#take a look
par(mfrow=c(2,2))
plot(SB.9)
plot(SB.9$residuals)

#place back in the data base with imputed data for SB's
pred.SB <- round(predict(SB.9, BB.df))
SB.imp.1 <- impute(BB.df$TEAM_BASERUN_SB, pred.SB)
BB.df$TEAM_BASERUN_SB <- SB.imp.1
```

#####Build model to replace DP
```{r}
DP <- lm(data=BB.df, TEAM_FIELDING_DP~.)
summary(DP)

#remove batting 2B's
DP.1 <- lm(data=BB.df, TEAM_FIELDING_DP~.-TEAM_BATTING_2B)
summary(DP.1)

#remove batting 3B's
DP.2 <- lm(data=BB.df, TEAM_FIELDING_DP~.-TEAM_BATTING_2B -TEAM_BATTING_3B)
summary(DP.2)

#remove batting 1B's
DP.3 <- lm(data=BB.df, TEAM_FIELDING_DP~.-TEAM_BATTING_2B -TEAM_BATTING_3B - TEAM_BATTING_1B)
summary(DP.3)

#remove all remaining batting
DP.4 <- lm(data=BB.df, TEAM_FIELDING_DP~.-TEAM_BATTING_2B -TEAM_BATTING_3B - TEAM_BATTING_1B - TEAM_BATTING_HR - TEAM_BATTING_SO - TEAM_BATTING_BB)
summary(DP.4)

#remove pitching strikeouts
DP.5 <- lm(data=BB.df, TEAM_FIELDING_DP~.-TEAM_BATTING_2B -TEAM_BATTING_3B - TEAM_BATTING_1B - TEAM_BATTING_HR - TEAM_BATTING_SO - TEAM_BATTING_BB - TEAM_PITCHING_SO)
summary(DP.5)


#all low P value and F statistic of 259.2 with adj R squared of 0.3936
#take a look
par(mfrow=c(2,2))
plot(DP.5)
plot(DP.5$residuals)

#place back in the data base with imputed data for SB's
pred.DP <- round(predict(DP.5, BB.df))
DP.imp.1 <- impute(BB.df$TEAM_FIELDING_DP, pred.DP)
BB.df$TEAM_FIELDING_DP <- DP.imp.1

#test new data set
summary(BB.df)
```


#Preliminary Transformation
####FIELDING ERRORS TRANSFORMATION
```{r trx1, eval=TRUE, echo=TRUE}
#most ever errors by a team is 639 by 1883 Philadelphia.  Prorating to 162 games gives a value of 1046.
BB.df$TEAM_FIELDING_E[which(BB.df$TEAM_FIELDING_E > 1046)] <- 159
```
####PITCHING HITS TRANSFORMATION
```{r trx2, eval=TRUE, echo=TRUE}
#most ever hits by a team is 1730.  So replace all pitching hits >3000 to be conservative with the median
BB.df$TEAM_PITCHING_H[which(BB.df$TEAM_PITCHING_H >3000)] <- 1518
```
####PITCHING STRIKEOUT TRANSFORMATION
The most strikeouts thrown in a single season (unadjusted) that I could find was fewer that 1400.
```{r trx3, eval=TRUE, echo=TRUE}
#REVISIT THIS WITH A LEVERAGE POINT TEST FOR BAD OUTLIERS!
#The TEAM_PITCHING_SO variable has 25 outliers that are far beyond the most team pitched strikeouts 
#that have ever occurred. Here's the line of R code to find the count:
nrow(data.frame(which(BB.df$TEAM_PITCHING_SO > 1450)))
#Since those outliers skew the distribution severely, I would suggest that rather than transform the 
#variable via a power transform we simply set those outliers to a more reasonable value, perhaps the 
#value of the 3rd quartile (968). This can be done (if we choose) with the following line of R code:
#replace pitching SO & pitching hits & errors outliers with median
BB.df$TEAM_PITCHING_SO[which(BB.df$TEAM_PITCHING_SO >1450)] <- 813
summary(BB.df)
```
####BATTING STRIKEOUT TRANSFORMATION
```{r trx4, eval=TRUE, echo=TRUE}
#First argument must be strictly positive.
#summary( powerTransform( cbind( TEAM_BATTING_SO, TEAM_BASERUN_SB, TEAM_PITCHING_SO, TEAM_FIELDING_DP) ~ 1, BB.df)
#This function call yields the "best" box-cox power transform exponent for each variable relative to the full data set #rather than limiting the results to being based solely on a single response/predictor pair.  The estimates are found in #the column labeled: "Est. Power". We need to round up or down to the nearest common transform as described here (reposted #from Blackboard discussion):
```
####BATTING SINGLE TRANSFORMATION
```{r trx5, eval=TRUE, echo=TRUE}
BB.df$TARGET_WINS[which(BB.df$TARGET_WINS <= 0)] <- 1
m1 <- lm(BB.df$TARGET_WINS~log(BB.df$TEAM_BATTING_1B))
#http://stats.stackexchange.com/questions/137059/find-distribution-and-transform-to-normal-distribution
lambda <- c(-1,-0.5, -0.33, -0.25, 0, 0.25, 0.33, 0.5,1)
#invResPlot(m1,lambda)
#inverseResponsePlot(m1,key=TRUE)
#lambda <- c(-1,-0.1811955,0,1)
#RSS <- c(6847.993,6761.037,6764.793,6901.701)
#plot(lambda,RSS,type="l",ylab=expression(RSS(lambda)),xlab=expression(lambda))
#-1/3
#ty <- y^(-1/3)
#plot(density(ty,kern="gaussian"),type="l",main="Gaussian kernel density estimate",xlab=expression(Y^(-1/3)))
#rug(ty)
```

# Correlations
```{r correlation, eval=TRUE, echo=FALSE}
par(cex = 0.65)
corrplot(
    cor(BB.df[c(1:13)]), 
    type='lower', 
    tl.srt=45,
    #tl.pos = "d",
    #is.corr=FALSE,
    #cl.lim=c(-1,1),
    addshade="positive",
    addCoef.col = rgb(0,0,0,alpha=0.3),
    addCoefasPercent = TRUE)
#par(cex = cex.before)
par(mfrow=c(1,1))
```
summary(BB.df)

Generalized Equation for Multiple Regression
\[
\begin{aligned}
\widehat{wins} &= \hat{\beta}_0 + 
                    \hat{\beta}_1 \times singles  + 
                    \hat{\beta}_2 \times doubles + 
                    \hat{\beta}_3 \times triples +
                    \hat{\beta}_4 \times homeruns + 
                    \hat{\beta}_5 \times walks + 
                    \hat{\beta}_6 \times strikeouts . . . +
                    \end{aligned}
\]

# Full Panel
```{r full, eval = TRUE,echo=TRUE}
#wins x full panel
pairs(TARGET_WINS~.,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
m1 <- lm(TARGET_WINS~., data = BB.df)
StanRes1 <- rstandard(m1)

summary(m1)

par(mfrow=c(2,2))
#SINGLES:                 transform skew before adding to model - high p-value
plot(density(BB.df$TEAM_BATTING_1B),main="Singles");rug(BB.df$TEAM_BATTING_1B)
plot(m1$model$TEAM_BATTING_1B,StanRes1,xlab="Singles",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BATTING_1B,StanRes1),lty=2,col=2)

#DOUBLES:                 add - high p-value
plot(density(BB.df$TEAM_BATTING_2B),main="Doubles");rug(BB.df$EAM_BATTING_2B)
plot(m1$model$TEAM_BATTING_2B,StanRes1,xlab="Doubles",ylab="Standardized 
Residuals");abline(lsfit(m1$model$TEAM_BATTING_2B,StanRes1),lty=2,col=2)

#TRIPLES:                 transform skew before adding to model - high p-value
plot(density(BB.df$TEAM_BATTING_3B),main="Triples");rug(BB.df$TEAM_BATTING_3B)
plot(m1$model$TEAM_BATTING_3B,StanRes1,xlab="Triples",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BATTING_3B,StanRes1),lty=2,col=2)

#HOMERUNS:                transform bimodal skew before adding to model - high p-value
plot(density(BB.df$TEAM_BATTING_HR),main="Homeruns");rug(BB.df$TEAM_BATTING_HR)
plot(m1$model$TEAM_BATTING_HR,StanRes1,xlab="Homeruns",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BATTING_HR,StanRes1),lty=2,col=2)

#WALKS:                   add
plot(density(BB.df$TEAM_BATTING_BB),main="Walks");rug(BB.df$TEAM_BATTING_BB)
plot(m1$model$TEAM_BATTING_BB,StanRes1,xlab="Walks",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BATTING_BB,StanRes1),lty=2,col=2)

#STRIKEOUTS:              transform bimodal skew before adding to model high p-value
plot(density(BB.df$TEAM_BATTING_SO),main="Strikeouts");rug(BB.df$TEAM_BATTING_SO)
plot(m1$model$TEAM_BATTING_SO,StanRes1,xlab="Strikeouts",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BATTING_SO,StanRes1),lty=2,col=2)

#HIT BY PITCH:            removed variable 
#plot(density(BB.df$TEAM_BATTING_HBP,na.rm=TRUE),main="Hit By Pitch");rug(BB.df$TEAM_BATTING_HBP)
#plot(m1$model$TEAM_BATTING_HBP,StanRes1,xlab="Singles",ylab="Standardized #Residuals");abline(lsfit(m1$model$TEAM_BATTING_HBP,StanRes1),lty=2,col=2)

#STOLEN BASES:            transform skew before adding to model 
plot(density(BB.df$TEAM_BASERUN_SB),main="Stolen Bases");rug(BB.df$TEAM_BASERUN_SB)
plot(m1$model$TEAM_BASERUN_SB,StanRes1,xlab="Stolen Bases",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_BASERUN_SB,StanRes1),lty=2,col=2)

#CAUGHT STEALING:         removed variable
#plot(density(BB.df$TEAM_BASERUN_CS),main="Caught Stealing");rug(BB.df$TEAM_BASERUN_CS)
#plot(m1$model$TEAM_BASERUN_CS,StanRes1,xlab="Caught Stealing",ylab="Standardized #Residuals");abline(lsfit(m1$model$TEAM_BASERUN_CS,StanRes1),lty=2,col=2)

#PITCHING HITS:           removed variable
#plot(density(BB.df$TEAM_PITCHING_H),main="Pitching Hits");rug(BB.df$TEAM_PITCHING_H)
#plot(m1$model$TEAM_PITCHING_H,StanRes1,xlab="Pitching Hits",ylab="Standardized #Residuals");abline(lsfit(m1$model$TEAM_PITCHING_H,StanRes1),lty=2,col=2)

#PITCHING HOMERUNS:       removed variable
#plot(density(BB.df$TEAM_PITCHING_HR),main="Pitching Homeruns");rug(BB.df$TEAM_PITCHING_HR)
#plot(m1$model$TEAM_PITCHING_HR,StanRes1,xlab="Pitching Homeruns",ylab="Standardized #Residuals");abline(lsfit(m1$model$TEAM_PITCHING_HR,StanRes1),lty=2,col=2)

#PITCHING WALKS:          correlated with batting walks
plot(density(BB.df$TEAM_PITCHING_BB),main="Pitching Walks");rug(BB.df$TEAM_PITCHING_BB)
plot(m1$model$TEAM_PITCHING_BB,StanRes1,xlab="Pitching Walks",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_PITCHING_BB,StanRes1),lty=2,col=2)

#PITCHING STRIKEOUTS:     add* - high p-value
plot(density(BB.df$TEAM_PITCHING_SO, na.rm = TRUE),main="Pitching Strikeouts");rug(BB.df$TEAM_PITCHING_SO)
plot(m1$model$TEAM_PITCHING_SO,StanRes1,xlab="Pitching Strikeouts",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_PITCHING_SO,StanRes1),lty=2,col=2)

#FIELDING ERRORS:         transform skew before adding to model   
plot(density(BB.df$TEAM_FIELDING_E),main="Fielding Errors");rug(BB.df$TEAM_FIELDING_E)
plot(m1$model$TEAM_FIELDING_E,StanRes1,xlab="Fielding Errors",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_FIELDING_E,StanRes1),lty=2,col=2)

#DOUBLE PLAYS:            add
plot(density(BB.df$TEAM_FIELDING_DP),main="Fielding Doubleplay");rug(BB.df$TEAM_FIELDING_DP)
plot(m1$model$TEAM_FIELDING_DP,StanRes1,xlab="Fielding Doubleplays",ylab="Standardized Residuals");abline(lsfit(m1$model$TEAM_FIELDING_DP,StanRes1),lty=2,col=2)
par(mfrow=c(1,1))
```

#Model Mo (Baseline) using forward selection
\[
\begin{aligned}
\widehat{wins} &= \hat{\beta}_0 + 
                    \hat{\beta}_1 \times doubles + 
                    \hat{\beta}_2 \times walks + 
                    \hat{\beta}_3 \times pitching strikeouts . . . +
                    \hat{\beta}_4 \times doubleplays                    
                    \end{aligned}
\]

#Model Mo
#Model M1
#Model M2
#Model M3
#Predictions and Assessment
#Conclusion
#Appendicies . . .
#### Slugging
The most hits in a single season (unadjusted) is 1783--NL Philadelphia Phillies 1930.
The most doubles in a single season (unadjusted) is 376--AL Texas in 2008.
The most triples in a single season (unadjusted) is 153--NL Baltimore in 1894.
Records that exceed these amounts should be adjusted either to NA or the median.

```{r summary-slugging, eval=FALSE, include=FALSE}
kable(DescStats.df[,c(15,2:4)],align='l',digits = c(0,2,0))
```

```{r boxplot-slugging, eval=FALSE, include=FALSE}
boxplot(BB.df[,3:6],main="Baseball Team Stats", 
        	xlab="Slugging") 
```

```{r slugging, eval=FALSE, include=FALSE}
#wins x batting - slugging
pairs(TARGET_WINS~TEAM_BATTING_1B+TEAM_BATTING_2B+TEAM_BATTING_3B+TEAM_BATTING_HR+TEAM_BATTING_BB+TEAM_BATTING_SO,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
m1 <- lm(TARGET_WINS~TEAM_BATTING_1B+TEAM_BATTING_2B+TEAM_BATTING_3B+TEAM_BATTING_HR+TEAM_BATTING_BB+TEAM_BATTING_SO
                  , data = BB.df)

par(mfrow=c(2,2))
plot(m1, ask=FALSE)

par(mfrow=c(3,3))
StanRes1 <- rstandard(m1)
plot(m1$model$TEAM_BATTING_1B,StanRes1,ylab="Standardized Residuals")
plot(m1$model$TEAM_BATTING_2B,StanRes1,ylab="Standardized Residuals")
plot(m1$model$TEAM_BATTING_3B,StanRes1,ylab="Standardized Residuals")
plot(m1$model$TEAM_BATTING_HR,StanRes1,ylab="Standardized Residuals")
plot(m1$model$TEAM_BATTING_BB,StanRes1,ylab="Standardized Residuals")
plot(m1$model$TEAM_BATTING_SO,StanRes1,ylab="Standardized Residuals")
plot(m1$fitted.values,StanRes1,ylab="Standardized Residuals",xlab="Fitted Values")

par(mfrow=c(3,3))
d <- density(m1$model$TARGET_WINS) # returns the density data
plot(d) # plots the results 
rug(m1$model$TARGET_WINS)

d <- density(m1$model$TEAM_BATTING_1B) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_1B)

d <- density(m1$model$TEAM_BATTING_2B) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_2B)

d <- density(m1$model$TEAM_BATTING_3B) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_3B)

d <- density(m1$model$TEAM_BATTING_HR) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_HR)

d <- density(m1$model$TEAM_BATTING_BB) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_BB)

d <- density(m1$model$TEAM_BATTING_SO) # returns the density data
plot(d) # plots the results 
rug(m1$model$TEAM_BATTING_SO)

par(mfrow=c(3,3))
avPlots(m1, ~.,ask=FALSE,identify.points=TRUE)
```

#### On-Base
102 strikeout NAs can remain as long as they're not counted in descriptive statistics 
as observations.  2085 hit by pitch NAs disqualifies this field from use in the model.

The most walks in a single season (unadjusted) is 835--AL Boston Red Sox in 1949.
The fewest walks in a single season (unadjusted) is 282--NL St. Louis Cardinals 1908.
Records that exceed these amounts should be adjusted either to NA or the median.

```{r summary-onbase, eval=FALSE, include=FALSE}
kable(DescStats.df[,c(5,6)],align='l',digits = c(0,2,0))
```

```{r boxplot-onbase, eval=FALSE, include=FALSE}
boxplot(BB.df[,c(5,6)],main="Baseball Team Stats", 
        	xlab="On Base") 
```

```{r onbase, eval=FALSE, include=FALSE}
#wins x batting - on-base
pairs(TARGET_WINS~TEAM_BATTING_BB+TEAM_BATTING_SO,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
ml.onbase <- lm(TARGET_WINS~TEAM_BATTING_BB+TEAM_BATTING_SO
                  , data = BB.df)
summary(ml.onbase)
```

#### Base Running
131 stolen base NAs can & 772 caught stealing NAs can remain as long as they're not counted
in descriptive statistics.

The most stolen bases in a single season (unadjusted) is 426--NL New York in 1893.
The most caught stealing bases count in a single season (unadjusted) is 191--AL NY in 1914.
Records that exceed these amounts should be adjusted either to NA or the median.

These statistics are co-linear and may be better used as a derived statistic for
Expected value of team stolen bases E(SB) = SB * likelihood of success (SB/SB attempts)

```{r summary-baserunning, eval=FALSE, include=FALSE}
kable(DescStats.df[,7,8],align='l',digits = c(0,2,0))
```

```{r boxplot-baserunning, eval =FALSE, include=FALSE}
boxplot(BB.df[,7,8],main="Baseball Team Stats", 
        	xlab="Base Running") 
```

```{r baserunning, eval=FALSE, include=FALSE}
#wins x batting - base running
pairs(TARGET_WINS~TEAM_BASERUN_SB+TEAM_BASERUN_CS,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
ml.baserunning <- lm(TARGET_WINS~TEAM_BASERUN_SB+TEAM_BASERUN_CS
                  , data = BB.df)
summary(ml.baserunning)
```

#### Fielding
286 Fielding error NAs can remain as long as they're not counted in descriptive statistics 
as observations.  

The most fielding errors in a single season (unadjusted) is 639--NL Philadelphia in 1883.
The most fielding errors in a single season (unadjusted) post WWII is 234--NL Philadelphia in 1945.
Records that exceed the post WWII amount should be adjusted either to NA or the median.

```{r summary-fielding, eval=FALSE, include=FALSE}
kable(DescStats.df[,13,14],align='l',digits = c(0,2,0))
```

```{r boxplot-fielding, eval=FALSE, include=FALSE}
boxplot(BB.df[,13,14],main="Baseball Team Stats", 
        	xlab="Fielding") 
```

```{r fielding, eval=FALSE, include=FALSE}
#wins x batting - base running
pairs(TARGET_WINS~TEAM_FIELDING_E+TEAM_FIELDING_DP,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
ml.fielding <- lm(TARGET_WINS~TEAM_FIELDING_E+TEAM_FIELDING_DP
                  , data = BB.df)
summary(ml.fielding)
```

#### Pitching
The most hits given up in a single season (unadjusted) that I could find was fewer that 2000.
The most homeruns given up in a single season (unadjusted) that I could find was fewer that 250.
The most walks given up in a single season (unadjusted) that I could find was fewer that 800.
The most strikeouts thrown in a single season (unadjusted) that I could find was fewer that 1400.
Records that exceed these amount should be adjusted either to NA or the median.

```{r summary-pitching, eval=FALSE, include=FALSE}
kable(DescStats.df[,9:12],align='l',digits = c(0,2,0))
```

```{r boxplot-pitching, eval = FALSE, include=FALSE}
boxplot(BB.df[,9:12],main="Baseball Team Stats", 
        	xlab="Pitching") 
```

```{r pitching, eval = FALSE, include=FALSE}
#wins x batting - pitching
pairs(TARGET_WINS~TEAM_PITCHING_H+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO,
        data=BB.df,pch=".",gap=.5,upper.panel=panel.smooth)
ml.pitching <- lm(TARGET_WINS~TEAM_PITCHING_H+TEAM_PITCHING_HR+TEAM_PITCHING_BB+TEAM_PITCHING_SO
                  , data = BB.df)
summary(ml.pitching)
```


